---
title: "DTFit_model_analysis_v2021-08-20"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls()) 
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(gridExtra)

#helper function #https://stackoverflow.com/questions/35775696/trying-to-use-dplyr-to-group-by-and-apply-scale
scale_this <- function(x) as.vector(scale(x))

# custom function to read a datatable
read_data <- function(directory, filename) {
  d = read.delim(paste(directory, filename, sep='/'), 
                 header=FALSE, sep='\t')
  
  #DATA EXCLUSION
  ## Exclude ngram model for having too many UNK tokens in EventsAdapt sentence set
  if (grepl("ngram", filename, fixed = TRUE) == TRUE) {
    message(">> Ngram model has too many UNK tokens for EventsAdapt sentence set! Excluding the model for now!")
    return(NULL)
  }
  ## Exclude GRNN & gpt2 surprisal model for only having been run on DTFit dataset
  if ((grepl("surprisal_scores_GRNN", filename, fixed = TRUE)  == TRUE) |
    (grepl("surprisal_scores_gpt2", filename, fixed = TRUE)  == TRUE)) {
    message(paste(">> Model", filename, "not run for EventsAdapt! Excluding the model!"))
    return(NULL)
  }
  
  #ROWS
  #set target number of *rows* in file
  target_rownumber = 798
  
  #check for target number of rows
  if (nrow(d)!=target_rownumber) {
    print(paste('unexpected number of sentences in file: ', filename, 'number of sentences: ', nrow(d)))
    return(NULL)
  }
  
  #COLUMNS
  #check for target number of *columns* in file
  if (ncol(d)==3 | ncol(d)==4 | ncol(d)==5 | ncol(d)==7) {
      
    target_trialnr = 2
    
    #streamline input format
    if (ncol(d)==3) {
      d = d  %>%
      rename(SentenceNum=V1, Sentence=V2, Score=V3)
    }
    else if (ncol(d)==4) {
      d = d  %>%
      select(V1,V2,V4) %>% #select relevant columns (do not choose Typicality column to streamline how it's assigned)
      rename(SentenceNum=V1, Sentence=V2, Score=V4)
    }
    
    else if (ncol(d)==5) { #If surprisal_scores_tinylstm
      d = d  %>%
      select(V1,V2,V5) %>% 
      rename(SentenceNum=V1, Sentence=V2, Score=V5)
    }
    
    else if (ncol(d)==7) {
      d = d  %>%
      select(V1,V3,V7) %>% #select relevant columns (others include grammatical tags)
      rename(SentenceNum=V1, Sentence=V3, Score=V7)
    }
    
    d = d %>%
      mutate(Typicality = ifelse(SentenceNum%%2==0, 'Atypical', 'Typical')) %>%
      mutate(ItemNum = SentenceNum %/% target_trialnr) %>%
      
      # 2. NAME MODEL_METRIC
      #add metric column
      mutate(Metric = substr(filename,1,nchar(filename)-4)) %>%
      
      #strip dataset prefix
      mutate(Metric = str_replace(Metric, "dtfit_vassallo_|dtfit.", "")) %>%
      
      ##Assimilate ppmi model names between datasets
      mutate(Metric = str_replace(Metric, "deps.smooth.baseline1", "syntax-based_PPMI")) %>%
      mutate(Metric = str_replace(Metric, "deps.smooth.baseline2", "order-based_PPMI")) %>%
      ##Assimilate gpt2 model names between datasets
      mutate(Metric = str_replace(Metric, "deps.prob_sent_gpt2-xl", "gpt2-xl_full")) %>%
      mutate(Metric = str_replace(Metric, "deps.prob_sent_gpt2", "gpt2_full"))
      
      # 3. PREPROCESS SCORES
    
      #3.1 Log-Scale GPT2 probabilities
      #the GPT2 models are the only ones with probability scores and not log probability! > log-scale
    if ((grepl("gpt2", filename) == TRUE) & (grepl("surprisal", filename) == FALSE)) {
      print(paste("log scaling scores in file: ", filename))
      d = d %>%
      mutate(Score = replace(Score, Score == 0.0, 1e-20)) %>% #Take out once 0.0 is fixed!
      mutate(Score = log(Score))
    }
      #3.2 Normalize scores for all models
    d = d %>%
      mutate(NormScore =scale_this(Score)) %>%

      # 4. PROCESS SENTENCES
      #strip space before final period
      mutate(Sentence = str_replace(Sentence, " [.]", ".")) %>%
      #for some reason, some baseline files have multiple periods at end > strip to just one!
      mutate(Sentence = str_replace(Sentence, "[.]{2,}", "."))
    
    return(d)
  } 
  else {
    print(paste('unexpected number of columns in file: ', 'number of columns: ', filename, ncol(d)))
    return(NULL)
  }
}
```

#SET ENVIRONMENT VARIABLES
```{r, echo=FALSE}
which_models <- "anns" #can be "anns", baselines" or "all"
message("Running with the following environment variables:")
message(paste("which_models: ", which_models))
```

## Below yields: 11172/798 = 14 models
```{r read data, echo=FALSE}
dirname = '../model_scores'

dirname_dtfit = paste(dirname, 'DTFit_scores_new_format', sep='/')

message("Loading DTFit data ...")
files = list.files(path=dirname_dtfit, pattern="*.txt")
dat = do.call(rbind, lapply(files, function(x) read_data(dirname_dtfit, x)))
```

###### PREPROCESS DATA FILES ######

## DTFIT
```{r read data, echo=FALSE}

dat$TrialType = 'DTFit'
```

### Filter out non-ANN models
```{r, echo=FALSE}

#print(unique(dat$Metric))
anns <- c("gpt2","bert","xlnet","roberta")

#filter out non-ANN models
if (which_models == "anns"){
  dat = dat %>%
  filter(grepl(paste(anns, collapse = "|"),dat$Metric))
  message("Using only ANN models for analysis")
} else if (which_models == "baselines"){
  dat = dat %>%
  filter(!grepl(paste(anns, collapse = "|"),dat$Metric))
  dat$Metric = factor(dat$Metric, levels=c("syntax-based_PPMI","order-based_PPMI","thematicFit_addition",
    "thematicFit_product","v2_sdm","fast_vector_sum","surprisal_scores_tinylstm"))
  message("Using only baseline models for analysis")
} else {
  message("Using all models for analysis")
}

print(unique(dat$Metric))
```

# TYPICALITY

We need to define for which models a higher score for typical sentences is desirable:

1. HIGHER = MORE TYPICAL
    * ANNs >> metric is probability
    * PPMI >> higher PMI indicates a typical sentence
    * Vector similarity
    * TFit similarity (sum of vector similarities to prototype vector)
    * SDM (sum of some vector similarities)
2. LOWER = MORE TYPICAL
    * Surprisal

```{r pressure, echo=FALSE}

lower_better = c("surprisal")
lower_better_pat <- paste(lower_better, collapse = '|')

dat.binchoice = dat %>%
  #create filter criterion
  mutate(LowerBetter = ifelse(grepl(lower_better_pat, Metric),TRUE,FALSE)) %>%
  group_by(ItemNum, Metric, LowerBetter, TrialType) %>%
  summarize(ScoreDiff = NormScore[Typicality=="Typical"]-NormScore[Typicality=="Atypical"]) %>%
  mutate(FinalScoreDiff = ifelse(LowerBetter==TRUE, -ScoreDiff, ScoreDiff)) %>%
  mutate(Accuracy = ifelse(FinalScoreDiff>0, 1, 0))

# Add FinalNormScore for NormScores plotting in the same direction for all metrics (i.e., Typical is more positive!)
# Take out/plot NormScore below if confusing!
dat = dat %>%
  mutate(LowerBetter = ifelse(grepl(lower_better_pat, Metric),TRUE,FALSE)) %>%
  mutate(FinalNormScore = ifelse(LowerBetter==TRUE, -NormScore, NormScore))
```

# BINARY ACCURACY

For baselines: Thematic fit models have lots & lots of 0.0 scores, which drags performance as Accuracy is set to 1 if the ScoreDifference is positive.
## Overall ANN plausibility plot
```{r, echo=FALSE}

ggplot(data=dat.binchoice, mapping=aes(x=TrialType,y=Accuracy))+
  stat_summary(geom='col', fun.y='mean',
               color='black', fill='gray80', width=0.5)+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', size = 0.5, width=0.1)+
  coord_cartesian(ylim=c(0,1))+
  theme_minimal()+
  xlab("")

savedir <- paste("plots/DTFit_Models=",which_models)
ifelse(!dir.exists(savedir), dir.create(savedir), FALSE)

savename <- "1_binaryAccuracy_all.png"
#print(paste(savedir,savename,sep="/"))

ggsave(paste(savedir,savename,sep="/"), width=16, height=12, units='cm')
```

# Split by model & metric
```{r, echo=FALSE, fig.height=12, fig.width=10}
ggplot(data=dat.binchoice, mapping=aes(x=TrialType, y=Accuracy))+
  facet_wrap(~Metric, ncol = 3)+
  stat_summary(geom='col', fun.y='mean',
               color='black', fill='gray80', width=0.5)+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', size = 0.5, width=0.1)+
  coord_cartesian(ylim=c(0,1))+
  theme_minimal()+
  xlab("")

savename <- "1_binaryAccuracy_split.png"
ggsave(paste(savedir,savename,sep="/"), width=16, height=12, units='cm')
```

# SCORE DIFFERENCE

Note: This is a weird plot for baselines as the score differences come from wildly different metrics!
Ignore the plots for baselines, or split up to enable a good visualization?
Also: As for some model a negative score difference indicates a more typical choice (e.g. surprisal models), and for some it indicates a more imtypical choice (e.g. PPMI models), I opted to plot the FinalScoreDiff.
FinalScoreDiff looks at whether a lower score is better for typical sentences and, if so, reverses the sign of the score difference (see above).
Note that we could also adjust the NormScore & reverse the sign, and then keep ScoreDiff. Haven't done that, because that is also not clean yet.

## Overall
```{r, echo=FALSE}
#ggplot(data=dat.binchoice, mapping=aes(x=TrialType, y=ScoreDiff))+
ggplot(data=dat.binchoice, mapping=aes(x=TrialType, y=FinalScoreDiff))+
  stat_summary(geom='col', fun.y='mean',
               color='black', fill='gray80', width=0.5)+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', size = 0.5, width=0.1)+
  theme_minimal()+
  xlab("")

savename <- "2_scoreDifference_all.png"
ggsave(paste(savedir,savename,sep="/"), width=16, height=12, units='cm')
```

##Split by model & metric
```{r, echo=FALSE, fig.height=12, fig.width=10}
#ggplot(data=dat.binchoice, mapping=aes(x=TrialType, y=ScoreDiff))+
ggplot(data=dat.binchoice, mapping=aes(x=TrialType, y=FinalScoreDiff))+
  facet_wrap(~Metric, ncol = 3)+
  stat_summary(geom='col', fun.y='mean',
               color='black', fill='gray80', width=0.5)+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', size = 0.5, width=0.1)+
  theme_minimal()+
  xlab("")

savename <- "2_scoreDifference_split.png"
ggsave(paste(savedir,savename,sep="/"), width=16, height=12, units='cm')
```


# SCORES by Typicality

Note that NormScore is more negative for typical sentences for some models (e.g. surprisal)
## Overall
```{r, echo=FALSE}
ggplot(data=dat, mapping=aes(x=TrialType, y=FinalNormScore, fill=Typicality))+
  stat_summary(geom='col', fun.y='mean',
               color='black', position='dodge', width=0.5)+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', position=position_dodge(width=0.5), size = 0.5, width=0.1)+
  xlab("")

savename <- "3_normScores_all.png"
ggsave(paste(savedir,savename,sep="/"), width=16, height=12, units='cm')
```

##Split by model & metric
```{r, echo=FALSE, fig.height=12, fig.width=10}

ggplot(data=dat, mapping=aes(x=TrialType, y=FinalNormScore, fill=Typicality))+
  facet_wrap(~Metric, ncol = 3)+
  stat_summary(geom='col', fun.y='mean',
               color='black', position='dodge', width=0.5)+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', position=position_dodge(width=0.5), size = 0.5, width=0.1)+
  xlab("")

savename <- "3_normScores_split.png"
ggsave(paste(savedir,savename,sep="/"), width=16, height=12, units='cm')
```