---
title: "model_analysis"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls()) 
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(gridExtra)

#helper function #https://stackoverflow.com/questions/35775696/trying-to-use-dplyr-to-group-by-and-apply-scale
scale_this <- function(x) as.vector(scale(x))

# custom function to read a datatable
read_data <- function(directory, filename) {
  d = read.delim(paste(directory, filename, sep='/'), 
                 header=FALSE, sep='\t')
  
  #DATA EXCLUSION
  ## Exclude ngram model for having too many UNK tokens in EventsAdapt sentence set
  if (grepl("ngram", filename, fixed = TRUE) == TRUE) {
    message(">> Ngram model has too many UNK tokens for EventsAdapt sentence set! Excluding the model for now!")
    return(NULL)
  }
  ## Exclude GRNN & gpt2 surprisal model for only having been run on EventsRev dataset
  if ((grepl("surprisal_scores_GRNN", filename, fixed = TRUE)  == TRUE) |
    (grepl("surprisal_scores_gpt2", filename, fixed = TRUE)  == TRUE)) {
    message(paste(">> Model", filename, "not run for EventsAdapt! Excluding the model!"))
    return(NULL)
  }
  
  #ROWS
  #set target number of *rows* in file
  if (grepl("Adapt", filename, fixed = TRUE) == TRUE) { #if EventsAdapt
    target_rownumber = 1648
  } else if (grepl("ev1", filename, fixed = TRUE) == TRUE){ #if EventsRev
    target_rownumber = 80
  } else {
    print(paste("unknown experiment for file: ", filename))
    target_rownumber = 0
  }
  #check for target number of rows
  if (nrow(d)!=target_rownumber) {
    print(paste('unexpected number of sentences in file: ', filename, 'number of sentences: ', nrow(d)))
    return(NULL)
  }
  
  #COLUMNS
  #check for target number of *columns* in file
  if (ncol(d)==3 | ncol(d)==4 | ncol(d)==6) {
      
      # 0. PREPARATION (set #of trials per item)
      if (grepl("Adapt", filename, fixed = TRUE) == TRUE) { #if EventsAdapt
        target_trialnr = 4
      } else if (grepl("ev1", filename, fixed = TRUE) == TRUE){ #if EventsRev
        target_trialnr = 2
      } else {
        print(paste("unknown experiment for file: ", filename))
        target_trialnr = 0
      }
    
    #streamline input format
    if (ncol(d)==3){
      d = d  %>%
      rename(SentenceNum=V1, Sentence=V2, Score=V3) %>%
      mutate(ItemNum = SentenceNum %/% target_trialnr)
    }
    else if (ncol(d)==4){
      d = d  %>%
      select(V1,V2,V4) %>% #select relevant columns (do not choose Plausibility column to streamline how it's assigned
                           #(e.g. all lowercase etc))
      rename(SentenceNum=V1, Sentence=V2, Score=V4) %>%
      mutate(ItemNum = SentenceNum %/% target_trialnr)
    }
    
    else if ((ncol(d)==6) & (grepl("smooth", filename) == FALSE)) { #If surprisal_scores_tinylstm
      #Add sentence number from 0 to len(dataset) -1
      tgt_len = nrow(d)-1
      sentnums = c(0:tgt_len)
      d = d  %>%
      select(V1,V5) %>% #select relevant columns (others include UNKification for EventsAdapt dataset)
      rename(Sentence=V1, Score=V5) %>%
      mutate(SentenceNum = sentnums) %>% #This adds a column SentenceNum from 0 to len(dataframe)-1
      mutate(ItemNum = SentenceNum %/% target_trialnr)
    }
    
    else if ((ncol(d)==6) & (grepl("smooth", filename) == TRUE)) {
      d = d  %>%
      select(V1,V3,V6) %>% #select relevant columns (others include grammatical tags)
      rename(SentenceNum=V1, Sentence=V3, Score=V6) %>%
      mutate(ItemNum = SentenceNum %/% target_trialnr)
    }
    
    d = d %>%
      mutate(Plausibility = ifelse(SentenceNum%%2==0, 'Plausible', 'Implausible')) %>%
      
      # 2. NAME MODEL_METRIC
      #add metric column
      mutate(Metric = substr(filename,1,nchar(filename)-4)) %>%
      
      #strip EventsAdapt dataset prefix
      mutate(Metric = str_replace(Metric, "new[_-][Ee]ventsAdapt[_\\.]", "")) %>%
      mutate(Metric = str_replace(Metric, "newsentences[_-]EventsAdapt[_\\.]", "")) %>%
      #strip EventsRev dataset prefix
      mutate(Metric = str_replace(Metric, "ev1[_\\.]", "")) %>%
      
      ##Assimilate ppmi model names between datasets
      mutate(Metric = str_replace(Metric, "deps.smooth.baseline1", "syntax-based_PPMI")) %>% #EventsRev
      mutate(Metric = str_replace(Metric, "scores_baseline1_ppmi", "syntax-based_PPMI")) %>% #EventsAdapt
      mutate(Metric = str_replace(Metric, "deps.smooth.baseline2", "order-based_PPMI")) %>% #EventsRev
      mutate(Metric = str_replace(Metric, "scores_baseline2_ppmi", "order-based_PPMI")) %>% #EventsAdapt
      ##Assimilate thematic fit model names between datasets
      mutate(Metric = str_replace(Metric, "update-model.TF-add", "thematicFit_addition")) %>%
      mutate(Metric = str_replace(Metric, "update-model.TF-prod", "thematicFit_product")) %>%
      ##Assimilate gpt2 model names between datasets
      mutate(Metric = str_replace(Metric, "deps.prob_sent_gpt2-xl", "gpt2-xl_full")) %>%
      mutate(Metric = str_replace(Metric, "deps.prob_sent_gpt2", "gpt2_full"))
      
      # 3. PREPROCESS SCORES
    
      #3.1 Log-Scale GPT2 probabilities
      #the GPT2 models are the only ones with probability scores and not log probability! > log-scale
      #Note: One entire quadruplet (SentenceNum 1516, The editor overheard the proofreader .) has score 0.0
      # >> add 1e-30 to enable log scaling & ask Emmanuele why this is 0!
    if ((grepl("gpt2", filename) == TRUE) & (grepl("surprisal", filename) == FALSE)) {
      print(paste("log scaling scores in file: ", filename))
      d = d %>%
      mutate(Score = replace(Score, Score == 0.0, 1e-20)) %>% #Take out once 0.0 is fixed!
      mutate(Score = log(Score))
    }
      #3.2 Normalize scores for all models
    d = d %>%
      mutate(NormScore =scale_this(Score)) %>%

      # 4. PROCESS SENTENCES
      #strip space before final period for alignment with TrialTypes etc below
      mutate(Sentence = str_replace(Sentence, " [.]", ".")) %>%
      #for some reason, some baseline files have multiple periods at end > strip to just one!
      mutate(Sentence = str_replace(Sentence, "[.]{2,}", "."))
    
    return(d)
  } 
  else {
    print(paste('unexpected number of columns in file: ', 'number of columns: ', filename, ncol(d)))
    return(NULL)
  }
}
```

#SET ENVIRONMENT VARIABLES
#use BOTH EventsAdapt AND EventsRev OR only EventsAdapt
```{r, echo=FALSE}
use_eventsrev <- TRUE #set to FALSE if you don't want to also use EventsRev data for analyses
which_models <- "anns" #can be "anns", baselines" or "all"
message("Running with the following environment variables:")
message(paste("use_eventsrev: ", use_eventsrev))
message(paste("which_models: ", which_models))
```

# Load data (below yields 28368 (lines)/1648 (sentences) = 16 (models)) for EventsAdapt
# Load data (below yields 960 (lines)/80 (sentences) = 12 (models)) for EventsRev
```{r read data, echo=FALSE}
dirname = '../model_scores'

dirname_eventsAdapt = paste(dirname, 'new-EventsAdapt', sep='/')
dirname_eventsRev = paste(dirname, 'EventsRev', sep='/')

message("Loading EventsAdapt data ...")
files_ea = list.files(path=dirname_eventsAdapt, pattern="*.txt")
dat_ea = do.call(rbind, lapply(files_ea, function(x) read_data(dirname_eventsAdapt, x)))

cat("\n")
message("Loading EventsRev data ...")
files_er = list.files(path=dirname_eventsRev, pattern="*.txt")
dat_er = do.call(rbind, lapply(files_er, function(x) read_data(dirname_eventsRev, x)))
```

###### PREPROCESS DATA FILES ######

## EVENTSADAPT
##Load & merge with original sentence file for information about TrialType & Voice
NOTE: Some of the sentences were different from the human experiment! >> take from dropbox, NOT behavioral study!
Maybe we should also just exclude these sentences
The file we use is downloaded from Dropbox and named here as 'newsentences_EventsAdapt_smallfixes.csv' to avoid confusion with the 'newsentences_EventsAdapt.csv' file in the behavioral study folder
```{r read data, echo=FALSE}
sentence_data = read.csv('newsentences_EventsAdapt_smallfixes.csv') %>%
rename(TrialType=EvType) %>%
rename(SentenceNum=X) %>%
rename(ItemNum=ItemNumber) %>% 
rename(Sentence=Stimulus) %>%
select(Sentence, Voice, TrialType)

#merge
dat_ea = merge(dat_ea, sentence_data)
dat_ea = arrange(dat_ea, desc(Metric), ItemNum, SentenceNum)

dat_ea$Experiment = 'EventsAdapt'

#align ItemNumber with numbers from behavioral analysis
dat_ea$ItemNum = dat_ea$ItemNum+1
```

## EVENTSREV
```{r read data, echo=FALSE}

dat_er$Experiment = 'EventsRev'
dat_er$TrialType = 'AAN'           # can change to a different value to show EventsRev results separately
dat_er$Voice = 'active'

#align ItemNumber with numbers from behavioral analysis
dat_er$ItemNum = dat_er$ItemNum + 1

dat_er$ItemNum = dat_er$ItemNum + 900    # so that the numbering doesn't overlap
```

## COMBINE OR NOT
```{r read data, echo=FALSE}
if (use_eventsrev == TRUE){
  dat = rbind(dat_er, dat_ea)
} else if (use_eventsrev == FALSE){
  dat = dat_ea
} else {
  stop("Please specify whether to use EventsRev for analysis or not!")
}

dat$TrialType = factor(dat$TrialType, levels=c('AI', 'AAN', 'AAR'))

# add info about synonym pairs for dat_ea items
dat = dat %>%
  mutate(SynonymPair = ifelse(ItemNum>900, NA, (ItemNum+1)%/%2)) %>%
  mutate(NumSyn = ifelse(ItemNum>900, NA, paste('Version', (ItemNum+1)%%2+1, sep="")))
```

### Filter out non-ANN models
### below yields 19908 (lines) /(1648+80=1728) (sentences) = 11 (models) for EventsAdapt + EventsRev
```{r, echo=FALSE}

#print(unique(dat$Metric))
anns <- c("gpt2","bert","xlnet","roberta")

#filter out non-ANN models
if (which_models == "anns"){
  dat = dat %>%
  filter(grepl(paste(anns, collapse = "|"),dat$Metric))
  message("Using only ANN models for analysis")
} else if (which_models == "baselines"){
  dat = dat %>%
  filter(!grepl(paste(anns, collapse = "|"),dat$Metric))
  dat$Metric = factor(dat$Metric, levels=c("syntax-based_PPMI","order-based_PPMI","thematicFit_addition",
    "thematicFit_product","v2_sdm","fast_vector_sum","surprisal_scores_tinylstm"))
  message("Using only baseline models for analysis")
} else {
  message("Using all models for analysis")
}

print(unique(dat$Metric))
```

# PLAUSIBILITY

We need to define for which models a higher score for plausible sentences is desirable:

1. HIGHER = MORE PLAUSIBLE
    * ANNs >> metric is probability
    * PPMI >> higher PMI indicates a plausible sentence
    * Vector similarity
    * TFit similarity (sum of vector similarities to prototype vector)
    * SDM (sum of some vector similarities)
2. LOWER = MORE PLAUSIBLE
    * Surprisal

```{r pressure, echo=FALSE}

lower_better = c("surprisal")
lower_better_pat <- paste(lower_better, collapse = '|')

dat.binchoice = dat %>%
  #create filter criterion
  mutate(LowerBetter = ifelse(grepl(lower_better_pat, Metric),TRUE,FALSE)) %>%
  group_by(ItemNum, TrialType, Voice, Metric,LowerBetter) %>%
  summarize(ScoreDiff = NormScore[Plausibility=="Plausible"]-NormScore[Plausibility=="Implausible"]) %>%
  mutate(FinalScoreDiff = ifelse(LowerBetter==TRUE, -ScoreDiff, ScoreDiff)) %>%
  mutate(Accuracy = ifelse(FinalScoreDiff>0, 1, 0))

# Add FinalNormScore for NormScores plotting in the same direction for all metrics (i.e., Plausible is more positive!)
# Take out/plot NormScore below if confusing!
dat = dat %>%
  mutate(LowerBetter = ifelse(grepl(lower_better_pat, Metric),TRUE,FALSE)) %>%
  mutate(FinalNormScore = ifelse(LowerBetter==TRUE, -NormScore, NormScore))
```

# BINARY ACCURACY

For baselines: Thematic fit models have lots & lots of 0.0 scores, which drags performance as Accuracy is set to 1 if the ScoreDifference is positive.
## Overall ANN plausibility plot
```{r, echo=FALSE}

ggplot(data=dat.binchoice, mapping=aes(x=TrialType, y=Accuracy))+
  stat_summary(geom='col', fun.y='mean',
               color='black', fill='gray80', width=0.5)+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', size = 0.5, width=0.1)+
  coord_cartesian(ylim=c(0,1))+
  theme_minimal()

savedir <- paste("plots/EventsRev=",use_eventsrev,"_Models=",which_models)
ifelse(!dir.exists(savedir), dir.create(savedir), FALSE)

savename <- "1_binaryAccuracy_all.png"
print(paste(savedir,savename,sep="/"))

ggsave(paste(savedir,savename,sep="/"), width=16, height=12, units='cm')
```

# Split by model & metric
```{r, echo=FALSE, fig.height=12, fig.width=15}
ggplot(data=dat.binchoice, mapping=aes(x=TrialType, y=Accuracy))+
  facet_wrap(~Metric, ncol = 3)+
  stat_summary(geom='col', fun.y='mean',
               color='black', fill='gray80', width=0.5)+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', size = 0.5, width=0.1)+
  coord_cartesian(ylim=c(0,1))+
  theme_minimal()

savename <- "1_binaryAccuracy_split.png"
ggsave(paste(savedir,savename,sep="/"), width=16, height=12, units='cm')
```


# SCORE DIFFERENCE

Note: This is a weird plot for baselines as the score differences come from wildly different metrics!
Ignore the plots for baselines, or split up to enable a good visualization?
Also: As for some models a negative score difference indicates a more plausible choice (e.g. surprisal models), and for some it indicates a more implausible choice (e.g. PPMI models), I opted to plot the FinalScoreDiff.
FinalScoreDiff looks at whether a lower score is better for plausible sentences and, if so, reverses the sign of the score difference (see above).
Note that we could also adjust the NormScore & reverse the sign, and then keep ScoreDiff. Haven't done that, because that is also not clean yet.

## Overall
```{r, echo=FALSE}
#ggplot(data=dat.binchoice, mapping=aes(x=TrialType, y=ScoreDiff))+
ggplot(data=dat.binchoice, mapping=aes(x=TrialType, y=FinalScoreDiff))+
  stat_summary(geom='col', fun.y='mean',
               color='black', fill='gray80', width=0.5)+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', size = 0.5, width=0.1)+
  theme_minimal()

savename <- "2_scoreDifference_all.png"
ggsave(paste(savedir,savename,sep="/"), width=16, height=12, units='cm')
```

##Split by model & metric
```{r, echo=FALSE, fig.height=15, fig.width=15}
#ggplot(data=dat.binchoice, mapping=aes(x=TrialType, y=ScoreDiff))+
ggplot(data=dat.binchoice, mapping=aes(x=TrialType, y=FinalScoreDiff))+
  facet_wrap(~Metric, ncol = 3)+
  stat_summary(geom='col', fun.y='mean',
               color='black', fill='gray80', width=0.5)+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', size = 0.5, width=0.1)+
  theme_minimal()

savename <- "2_scoreDifference_split.png"
ggsave(paste(savedir,savename,sep="/"), width=30, height=30, units='cm')
```


# SCORES by Plausibility

Note that NormScore is more negative for plausible sentences for some models (e.g. surprisal)
## Overall
```{r, echo=FALSE}
ggplot(data=dat, mapping=aes(x=TrialType, y=FinalNormScore, fill=Plausibility))+
  stat_summary(geom='col', fun.y='mean',
               color='black', position='dodge', width=0.5)+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', position=position_dodge(width=0.5), size = 0.5, width=0.1)

savename <- "3_normScores_all.png"
ggsave(paste(savedir,savename,sep="/"), width=16, height=12, units='cm')
```

##Split by model & metric
```{r, echo=FALSE, fig.height=12, fig.width=15}

ggplot(data=dat, mapping=aes(x=TrialType, y=FinalNormScore, fill=Plausibility))+
  facet_wrap(~Metric, ncol = 3)+
  stat_summary(geom='col', fun.y='mean',
               color='black', position='dodge', width=0.5)+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', position=position_dodge(width=0.5), size = 0.5, width=0.1)

savename <- "3_normScores_split.png"
ggsave(paste(savedir,savename,sep="/"), width=16, height=12, units='cm')
```

# Within-item scores for plausible vs. implausible versions of the sentence

## Overall
```{r, echo=FALSE}
dat.wide = dat %>% 
  dplyr::select(ItemNum, TrialType, Plausibility, Voice, Metric, FinalNormScore) %>%
  group_by(ItemNum, TrialType, Plausibility, Voice, Metric) %>%
  summarize(meanScore = mean(FinalNormScore, na.rm=TRUE)) %>%
  spread(Plausibility, meanScore)  

ggplot(data=dat.wide)+
  geom_point(mapping=aes(x=Implausible, y=Plausible))+
  facet_wrap(~TrialType)+
  geom_abline()+
  theme_classic()

savename <- "4_withinItem_all.png"
ggsave(paste(savedir,savename,sep="/"), width=16, height=12, units='cm')
```

##Split by model & metric

```{r Fig1, echo=FALSE, fig.height=15, fig.width=15}
models_metrics = unique(dat$Metric)

for(i in seq_along(models_metrics)){
  nam <- paste("A", i, sep = "")
  #print(models_metrics[i])
  
  curr_data = dat.wide %>%
    filter(Metric==models_metrics[i])
  
  p <- ggplot(data=curr_data)+
  geom_point(mapping=aes(x=Implausible, y=Plausible))+
  facet_wrap(~TrialType)+
  geom_abline()+
  theme_classic()+
  ggtitle(paste(models_metrics[i]))
  
  assign(nam, p)
}

if (which_models=="all") {
  grid.arrange(A1, A2, A3, A4, A5, A6, A7, A8, A9, A10, A11, A12, A13, A14, A15, A16, A17, A18)
  g <- arrangeGrob(A1, A2, A3, A4, A5, A6, A7, A8, A9, A10, A11, A12, A13, A14, A15, A16, A17, A18) #generates g
} else if (which_models=="anns") {
  grid.arrange(A1, A2, A3, A4, A5, A6, A7, A8, A9, A10, A11)
  g <- arrangeGrob(A1, A2, A3, A4, A5, A6, A7, A8, A9, A10, A11)
} else { #if which_models == "baselines"
  grid.arrange(A1, A2, A3, A4, A5, A6, A7)
  g <- arrangeGrob(A1, A2, A3, A4, A5, A6, A7)
  }
savename <- "4_withinItem_split.png"
ggsave(paste(savedir,savename,sep="/"), width=30, height=30, units='cm', g)
```

# ACTIVE VS PASSIVE

## Overall

```{r, echo=FALSE}
#ggplot(data=dat.binchoice, mapping=aes(x=TrialType, y=ScoreDiff, fill=Voice))+
ggplot(data=dat.binchoice, mapping=aes(x=TrialType, y=FinalScoreDiff, fill=Voice))+
  stat_summary(geom='col', fun.y='mean', position='dodge', width=0.5)+
  stat_summary(geom='errorbar', fun.data='mean_se', position=position_dodge(width=0.5),
               color = 'black', size = 0.5, width=0.1)

savename <- "5_activePassive_all.png"
ggsave(paste(savedir,savename,sep="/"), width=16, height=12, units='cm')
```

##Split by model & metric

```{r, echo=FALSE, fig.height=12, fig.width=15}
#ggplot(data=dat.binchoice, mapping=aes(x=TrialType, y=ScoreDiff, fill=Voice))+
ggplot(data=dat.binchoice, mapping=aes(x=TrialType, y=FinalScoreDiff, fill=Voice))+
  facet_wrap(~Metric, ncol = 3)+
  stat_summary(geom='col', fun.y='mean', position='dodge', width=0.5)+
  stat_summary(geom='errorbar', fun.data='mean_se', position=position_dodge(width=0.5),
               color = 'black', size = 0.5, width=0.1)

savename <- "5_activePassive_split.png"
ggsave(paste(savedir,savename,sep="/"), width=16, height=12, units='cm')
```


## By item

```{r, echo=FALSE}
dat.voice = dat %>%
  filter(!is.na(SynonymPair)) %>%
  dplyr::select(ItemNum, TrialType, Plausibility, Voice, Metric, FinalNormScore) %>%
  group_by(ItemNum, TrialType, Plausibility, Voice, Metric) %>%
  summarize(meanScore = mean(FinalNormScore, na.rm=TRUE)) %>%
  spread(Voice, meanScore)

ggplot(data=dat.voice)+
  geom_point(mapping=aes(x=active, y=passive))+
  facet_wrap(~TrialType)+
  geom_abline()+
  theme_classic()

savename <- "6_activePassive_byitem_all.png"
ggsave(paste(savedir,savename,sep="/"), width=16, height=12, units='cm')
```

```{r Fig2, echo=FALSE, fig.height=15, fig.width=15}
models_metrics = unique(dat$Metric)

for(i in seq_along(models_metrics)){
  nam <- paste("A", i, sep = "")
  print(models_metrics[i])
  
  curr_data = dat.voice %>%
    filter(Metric==models_metrics[i])
  
  p <- ggplot(data=curr_data)+
  geom_point(mapping=aes(x=active, y=passive))+
  facet_wrap(~TrialType)+
  geom_abline()+
  theme_classic()+
  ggtitle(paste(models_metrics[i]))
  
  assign(nam, p)
}

if (which_models=="all") {
  grid.arrange(A1, A2, A3, A4, A5, A6, A7, A8, A9, A10, A11, A12, A13, A14, A15, A16, A17, A18)
  g <- arrangeGrob(A1, A2, A3, A4, A5, A6, A7, A8, A9, A10, A11, A12, A13, A14, A15, A16, A17, A18) #generates g
} else if (which_models=="anns") {
  grid.arrange(A1, A2, A3, A4, A5, A6, A7, A8, A9, A10, A11)
  g <- arrangeGrob(A1, A2, A3, A4, A5, A6, A7, A8, A9, A10, A11)
} else { #if which_models == "baselines"
  grid.arrange(A1, A2, A3, A4, A5, A6, A7)
  g <- arrangeGrob(A1, A2, A3, A4, A5, A6, A7)
  }
savename <- "6_activePassive_byitem_split.png"
ggsave(paste(savedir,savename,sep="/"), width=30, height=30, units='cm', g)
```

# SYNONYMS

## By item

```{r, echo=FALSE}
dat.syn = dat %>%
  filter(!is.na(SynonymPair)) %>%
  dplyr::select(-ItemNum) %>%
  group_by(TrialType, NumSyn, SynonymPair, Plausibility, Voice, Metric) %>%
  summarize(meanScore = mean(FinalNormScore, na.rm=TRUE)) %>%
  spread(NumSyn, meanScore)

ggplot(data=dat.syn)+
  geom_point(mapping=aes(x=Version1, y=Version2, color=Voice))+
  facet_wrap(~TrialType)+
  geom_abline()+
  theme_classic()

savename <- "7_synonyms_all.png"
ggsave(paste(savedir,savename,sep="/"), width=16, height=12, units='cm')
```

```{r, echo=FALSE, fig.height=15, fig.width=15}
models_metrics = unique(dat$Metric)

for(i in seq_along(models_metrics)){
  nam <- paste("A", i, sep = "")
  print(models_metrics[i])
  
  curr_data = dat.syn %>%
    filter(Metric==models_metrics[i])
  
  p <- ggplot(data=curr_data)+
  geom_point(mapping=aes(x=Version1, y=Version2, color=Voice))+
  facet_wrap(~TrialType)+
  geom_abline()+
  theme_classic()+
  ggtitle(paste(models_metrics[i]))
  
  assign(nam, p)
}

if (which_models=="all") {
  grid.arrange(A1, A2, A3, A4, A5, A6, A7, A8, A9, A10, A11, A12, A13, A14, A15, A16, A17, A18)
  g <- arrangeGrob(A1, A2, A3, A4, A5, A6, A7, A8, A9, A10, A11, A12, A13, A14, A15, A16, A17, A18) #generates g
} else if (which_models=="anns") {
  grid.arrange(A1, A2, A3, A4, A5, A6, A7, A8, A9, A10, A11)
  g <- arrangeGrob(A1, A2, A3, A4, A5, A6, A7, A8, A9, A10, A11)
} else { #if which_models == "baselines"
  grid.arrange(A1, A2, A3, A4, A5, A6, A7)
  g <- arrangeGrob(A1, A2, A3, A4, A5, A6, A7)
  }
savename <- "7_synonyms_split.png"
ggsave(paste(savedir,savename,sep="/"), width=30, height=30, units='cm', g)
```


