---
title: "DTFit_model_analysis_v2021-08-20"
output: html_document
---

# SETUP

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls()) 
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(gridExtra)
library(cocor)
library(patchwork)
library(lme4)
library(lmerTest)
library(gtools)

source('dataloader_utils.R') #includes normalizations, read_data functions
source('stats_utils.R')

# Suppress summarise info
options(dplyr.summarise.inform = FALSE)
```

### Print environment variables 
```{r}
#SET ENVIRONMENT VARIABLES
which_models <- "all_main" #can be "anns_all", "anns_main", or "all_main"
# anns_main = human + main ANNs
# all_main = human + main ANNs + main baselines
# anns_all = human + all ANNs
normalization <- "min-max" #can be "min-max", "zscore"

path <- paste("results/")
ifelse(!dir.exists(path), dir.create(path), FALSE)
savedir <- paste(path,"DTFit_Models=",which_models,sep='')
ifelse(!dir.exists(savedir), dir.create(savedir), FALSE)

print(paste(savedir))
message("Running with the following environment variables:")
message(paste("which_models: ", which_models))
message(paste("normalization: ", normalization))
```
```{r}
if (grepl("min-max", normalization) == TRUE){
  normalization <- function(x) {
    return(min_max(x)) }
  reference_value = 0.5
} else {
  normalization <- function(x) {
    return(scale_this(x)) }
  reference_value = 0
}
```

# READ DATA
```{r}
#created via preprocess_scores.Rmd
dat = read.csv('clean_data/clean_DTFit_df.csv')
```

# Select which models/metrics to plot
```{r, utils.choose_models, echo=FALSE}
human <- c("human")
anns <- c("GPT2","BERT","RoBERTa","tinyLSTM")
anns_main <- c("GPT2-XL.l2r","BERT-large.plast","RoBERTa-large.plast","tinyLSTM.surprisal") #best metrics
baselines <- c("syntax.PPMI","SDM")

anns_main <- c(human, anns_main)
all_main <- c(anns_main, baselines)

unique_metrics = unique(dat$Metric)
anns <- unique_metrics[grepl(paste(anns, collapse = "|"), unique_metrics)]
anns = mixedsort(anns)

anns_all <- anns

# select which to use
chosen_models = eval(parse(text=which_models))
dat = dat %>% filter(Metric %in% chosen_models)

message("Using these models/metrics:")
dat$Metric = factor(dat$Metric, levels = chosen_models)
print(unique(dat$Metric))
```


# CONTRASTS

```{r}
dat$Plausibility = factor(dat$Plausibility, levels=c("Plausible", "Implausible")) # dummy coding by default
dat$ItemNum = as.factor(dat$ItemNum)
```


# BINARY ACCURACY

## General plotting settings

```{r}
#Set plotting options for grid plots
nr_models = length(chosen_models)
ncols=nr_models
nrows=round(nr_models/ncols)
if (nrows * ncols < nr_models) {
  nrows = nrows + 1
}
```

```{r}

lower_better = c("surprisal")
lower_better_pat <- paste(lower_better, collapse = '|')

#add Category color
if (which_models == "anns_all") {
  dat.binchoice = dat %>%
  mutate(Metric = as.character(Metric)) %>%
  mutate(Category = ifelse(startsWith(Metric, "BERT"), "BERT", "RoBERTa")) %>%
  mutate(Category = ifelse(startsWith(Metric, "GPT2"), "GPT2", Category)) %>%
  mutate(Category = ifelse(startsWith(Metric, "tiny"), "tinyLSTM", Category)) %>%
  mutate(Category = ifelse(Metric=="human", "human", Category))
} else {
  dat.binchoice = dat %>%
  mutate(Category = ifelse(Metric%in%anns, "ANNs", "baselines")) %>%
  mutate(Category = ifelse(Metric=="human", "human", Category))
}

dat.binchoice = dat.binchoice %>%
  mutate(LowerBetter = ifelse(grepl(lower_better_pat, Metric),TRUE,FALSE)) %>%
  group_by(ItemNum, TrialType, Metric, LowerBetter, Category) %>%
  summarize(ScoreDiff = FinalNormScore[Plausibility=="Plausible"]-FinalNormScore[Plausibility=="Implausible"]) %>%
  mutate(FinalScoreDiff = ifelse(LowerBetter==TRUE, -ScoreDiff, ScoreDiff)) %>%
  mutate(Accuracy = ifelse(FinalScoreDiff>0, 1, 0)) %>%
  ungroup()

# specify order for plotting
dat.binchoice$Metric = factor(dat.binchoice$Metric, levels=chosen_models)

if (which_models=="anns_all") {
  dat.binchoice$Category = factor(dat.binchoice$Category, levels=c("human", "GPT2", "BERT", "RoBERTa", "tinyLSTM"))
} else {
  dat.binchoice$Category = factor(dat.binchoice$Category, levels=c("human", "ANNs", "baselines"))
}
```

### Stats

```{r}
# get p values
dat.binchoice.summary = dat.binchoice %>%
  group_by(Category, Metric) %>%
  summarize(NumCorrect=sum(Accuracy), NumTotal=length(Accuracy)) %>%
  mutate(AccuracyScore = NumCorrect/NumTotal) %>%
  ungroup() %>%
  mutate(pVal = calculate_binom_pval(NumCorrect, NumTotal))

# adjust for multiple comparisons within each category
dat.binchoice.summary = dat.binchoice.summary %>%
  group_by(Category) %>%
  mutate(pValAdjusted = p.adjust(pVal, method="fdr", n=length(pVal)),
         ntoadjust = length(pVal)) %>%
  mutate(pLabel = ifelse(pValAdjusted<0.001, "***", 
                         ifelse(pValAdjusted<0.01, "**",
                                ifelse(pValAdjusted<0.05, "*", "")))) %>%
  ungroup()
```


```{r}
# get human responses separately
human.results = dat.binchoice.summary %>%
  filter(Metric=='human') %>%
  select(NumCorrect, NumTotal) %>%
  rename(NumCorrectHuman=NumCorrect, NumTotalHuman=NumTotal)
dat.binchoice.summary.withchisq = merge(dat.binchoice.summary, human.results)
dat.binchoice.summary.withchisq = dat.binchoice.summary.withchisq %>%
  mutate(ChiSq = calculate_chisq_vectorized_chi(NumCorrect, NumTotal, NumCorrectHuman),
         pVal2humans = calculate_chisq_vectorized_p(NumCorrect, NumTotal, NumCorrectHuman)) %>%
  group_by(Category) %>%
  mutate(pVal2humansAdjusted = p.adjust(pVal2humans, method="fdr", n=length(pVal2humans)),
         ntoadjust = length(pVal2humans)) %>%
  mutate(pLabel2humans = ifelse(pVal2humansAdjusted<0.001, "***", 
                         ifelse(pVal2humansAdjusted<0.01, "**",
                                ifelse(pVal2humansAdjusted<0.05, "*", "")))) 
```

### Plot

```{r, echo=FALSE}

plot.binacc = ggplot(data=dat.binchoice, 
       mapping=aes(x=Metric, y=Accuracy, fill=Category))+
  geom_hline(yintercept=1, color='gray50', linetype='dotted')+
  stat_summary(geom='col', fun.y='mean',
               color='black', width=0.8)+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', size = 0.5, width=0.1)+
  geom_text(mapping=aes(x=Metric, y=0.05, label=pLabel), data=dat.binchoice.summary)+
  coord_cartesian(ylim=c(0,1))+
  geom_hline(yintercept=.5, linetype='dotted')+
  theme_classic()+
  labs(x='Model')+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
plot.binacc


savename <- "1_binaryAccuracy.png"
ggsave(paste(savedir,savename,sep="/"), width=15, height=9, units='cm')
```

## Error analysis

```{r}
acc_by_sentence.ANN = dat.binchoice %>% 
  filter(Category=='ANNs') %>%
  group_by(ItemNum) %>%
  mutate(meanAccuracy=mean(Accuracy))
```

### Quick plot
```{r}
ggplot(acc_by_sentence.ANN, aes(x=meanAccuracy)) + geom_histogram()
```

### Save hard sentences

```{r}
hard_sents = acc_by_sentence.ANN %>%
  filter(meanAccuracy<0.5) %>%
  select(ItemNum, meanAccuracy)

hard_sents = inner_join(hard_sents, dat) %>%
  select(ItemNum, Plausibility, Sentence, meanAccuracy) %>% 
  distinct() %>%
  spread(Plausibility, Sentence)

savename <- "hard_sentences_ANNs.csv"
write.csv(hard_sents, paste(savedir,savename,sep="/"), row.names=FALSE)
```

### Compare with human scores

```{r}
dat.binchoice.human = dat.binchoice %>% filter(Metric=='human')
dat.binchoice.human = inner_join(dat.binchoice.human, acc_by_sentence.ANN %>% select(ItemNum, meanAccuracy)) %>%
  distinct()

ggplot(dat.binchoice.human, aes(x=meanAccuracy, y=FinalScoreDiff)) + 
  stat_summary(geom='col', fun.y='mean')+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', size = 0.5, width=0.1)
```

TODO: regression stats

```{r}
ggplot(dat.binchoice.human, aes(x=meanAccuracy, y=FinalScoreDiff)) + 
  geom_jitter(width=0.02)

cor(dat.binchoice.human$meanAccuracy, dat.binchoice.human$FinalScoreDiff)
```

# PLAUSIBILITY
```{r, echo=FALSE}
# color scheme
color_plaus = '#1b9e77'
color_voice = '#d95f02'
color_syn = '#7570b3'
cortext_size = 3

dat.plaus = dat %>% 
  group_by(ItemNum, TrialType, Plausibility, Metric) %>%
  summarize(meanScore = mean(FinalNormScore, na.rm=TRUE)) %>%
  spread(Plausibility, meanScore)  

dat.plaus = dat.plaus %>%
  filter(Metric%in%chosen_models)
```

### Stats
```{r}
df_correlation = get_correlation_df("plausibility", "human", dat.plaus, chosen_models)
```

### Plot
```{r}
df_correlation$Metric = factor(df_correlation$Metric, levels=chosen_models)
dat.plaus$Metric = factor(dat.plaus$Metric, levels=chosen_models)

plot.plaus_anns = ggplot(data=subset(dat.plaus))+
  geom_point(mapping=aes(x=Plausible, y=Implausible), size=0.2, color=color_plaus)+
  facet_wrap(~Metric, ncol=ncols, nrow=nrows)+
  geom_abline(slope=1, intercept=0, size=0.2)+
  geom_point(mapping=aes(x=Plausible, y=Implausible), size=0.2, color=color_plaus)+
  geom_text(mapping=aes(x=0.05, y=1, label=sprintf("r = %.2f%s", round(Correlation,2), pVal2zeroLabel)), 
            data=df_correlation, size=cortext_size, hjust = 0)+
  coord_cartesian(ylim=c(0,1), xlim=c(0,1))+
  theme_classic()+
  theme(panel.spacing.x = unit(1, "lines"))+
  theme(plot.title = element_text(hjust=0.5,face='bold'))
plot.plaus_anns

savename <- "2_plausibility_scatter_anns.png"
ggsave(paste(savedir,savename,sep="/"), width=20, height=8, units='cm')
```
```{r, fig.height=2,fig.width=5}

df_correlation$Metric = factor(df_correlation$Metric, levels=chosen_models)
dat$Metric = factor(dat$Metric, levels=chosen_models)

plot.plaus.density.sub = ggplot()+
  geom_density(data=dat, mapping=aes(FinalNormScore, fill = Plausibility), alpha = 0.2)+
  facet_wrap(~Metric,nrow=nrows, ncol=ncols)+
  geom_text(mapping=aes(x=0.1, y=5, label=sprintf("r = %.2f%s", round(Correlation,2), pVal2zeroLabel)),  data=df_correlation, size=cortext_size, hjust = 0)+
  #coord_flip()+
  theme_classic()+
  xlab("Normed score") +
  ylab("Density")+
  theme(legend.title = element_blank())
plot.plaus.density.sub
```

# MODEL HUMAN COMPARISON

## Prep data

```{r}
dat.human = dat %>% filter(Metric=="human") %>%
  select(ItemNum, TrialType, Sentence, Plausibility, FinalNormScore)
dat.model = dat %>% filter(Metric!="human") %>%
  select(ItemNum, TrialType, Sentence, Plausibility, FinalNormScore, Metric)
dat.model2human = merge(dat.human, dat.model, 
                        by=c("ItemNum", "TrialType", "Sentence", "Plausibility")) %>%
  rename(HumanScore=FinalNormScore.x, ModelScore=FinalNormScore.y)
```

Formatting:

```{r, echo=FALSE, fig.height=12, fig.width=15}
dat.model2human = dat.model2human %>%
  #add color info
  mutate(Category = ifelse(Metric%in%anns, "ANNs", "baselines")) 

# specify order for plotting
dat.model2human$Metric = factor(dat.model2human$Metric, levels=chosen_models)
dat.model2human$Category = factor(dat.model2human$Category, levels=c("human", "ANNs", "baselines"))
```

## Calculate correlations for the plot

Cannot do a human2human comparison because we only have mean data

```{r}
df_correlation = data.frame()
metrics = unique(dat.model2human$Metric)

trialTypes = unique(dat.model2human$TrialType)
for (t in seq_along(trialTypes)){
  dat.model2human.tt = dat.model2human %>%
    filter(TrialType==trialTypes[t])
  
  for(i in seq_along(metrics)){
      current_metric=metrics[i]
      val1.model = dat.model2human.tt$ModelScore[dat.model2human.tt$Metric==current_metric]
      val2.model = dat.model2human.tt$HumanScore[dat.model2human.tt$Metric==current_metric]
      pval2zero = cor.test(val1.model, val2.model, method="pearson")$p.value
      corval = cor(val1.model, val2.model, method="pearson")
      
      # add vector to a dataframe
      df <- data.frame(current_metric, trialTypes[t], corval, pval2zero)
      df_correlation <- rbind(df_correlation,df)
  }
}
colnames(df_correlation) = c("Metric", "TrialType", "Correlation", "pVal2zero")

# adjust for multiple comparisons
df_correlation = df_correlation %>%
  mutate(pVal2zeroAdjusted = p.adjust(pVal2zero, method="fdr", n=length(pVal2zero))) %>%
  mutate(pVal2zeroLabel = ifelse(pVal2zeroAdjusted<0.001, "***", 
                                 ifelse(pVal2zeroAdjusted<0.01, "**",
                                        ifelse(pVal2zeroAdjusted<0.05, "*", ""))))

```

## Plot
```{r}

# text size 
cortext_size = 3

plot.model2human = ggplot(dat=dat.model2human)+  
  geom_abline(slope=1, intercept=0, size=0.2)+
  geom_point(mapping=aes(x=ModelScore, y=HumanScore), 
             size=0.4, position=position_jitter(width=0.01), color='#00BFC4')+
  facet_wrap(~Metric, ncol=ncols)+
  geom_text(mapping=aes(x=0.05, y=1.05, 
                        label=sprintf("r = %.2f%s", round(Correlation,2), pVal2zeroLabel)), 
            data=df_correlation, size=cortext_size, hjust = 0)+
  coord_cartesian(ylim=c(0,1.05), xlim=c(0,1))+
  scale_x_continuous(breaks=c(0,0.5,1))+
  scale_y_continuous(breaks=c(0,0.5,1))+
  theme_classic()+
  theme(panel.spacing.x = unit(1, "lines"))+
  theme(legend.position="none")
plot.model2human

#savename <- "model2human_byTrialType_ANNs.png"
#ggsave(paste(savedir,savename,sep="/"), height=15,width=25, units='cm')
```

# COMBINE

```{r}
(plot.binacc | plot_spacer() | 
   (((plot_spacer() + plot.model2human) + plot_layout(widths=c(1,8))) / plot.plaus_anns) ) + 
  plot_layout(widths=c(1.5,0.05, 3))

if (which_models == 'anns_all') {
  height = 15
  width = 55
} else if (which_models == 'all_main') {
  height = 15
  width = 40
} else {
  height = 15
  width = 30
}

savename <- "composite.png"
ggsave(paste(savedir,savename,sep="/"), height=height,width=width, units='cm')
```

```{r}

(plot.binacc | plot_spacer() | 
   (((plot_spacer() + plot.model2human) + plot_layout(widths=c(1,8))) / plot.plaus.density.sub) ) + 
  plot_layout(widths=c(1.5,0.05, 3))

savename <- "composite_density.png"
ggsave(paste(savedir,savename,sep="/"), height=height,width=width, units='cm')
```

# SCORE DIFFERENCE
```{r, echo=FALSE}

df_correlation = data.frame()
# get p values for t test
for(i in seq_along(chosen_models)){
  curr_dat = subset(dat.binchoice,Metric==chosen_models[i])
  curr_score_diff = c(curr_dat$FinalScoreDiff)
  print(paste(t.test(curr_score_diff)$p.value))
  ttestPval2zero = t.test(curr_score_diff)$p.value
  print(paste(chosen_models[i], ttestPval2zero))
  curr_cat = unique(curr_dat$Category)
  # add vector to a dataframe
    df <- data.frame(chosen_models[i], ttestPval2zero,curr_cat)
    df_correlation <- rbind(df_correlation,df)
}

# adjust for multiple comparisons within each category
df_correlation = df_correlation %>%
  mutate(pLabel = ifelse(ttestPval2zero<0.001, "***", 
                         ifelse(ttestPval2zero<0.01, "**",
                                ifelse(ttestPval2zero<0.05, "*",
                                       ifelse(ttestPval2zero>=0.05, "n.s.", "")))))

colnames(df_correlation) = c("Metric", "ttestPval2zero", "Category", "pLabel")

df_correlation$Metric = factor(df_correlation$Metric, levels=chosen_models)
dat.binchoice$Metric = factor(dat.binchoice$Metric, levels=chosen_models)

plot.score_difference = ggplot(data=dat.binchoice,mapping=aes(x=TrialType, y=FinalScoreDiff, fill=Category))+
  facet_wrap(~Metric, ncol=ncols, nrow=nrows)+
  stat_summary(geom='col', fun.y='mean',
               color='black', width=0.5)+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', size = 0.5, width=0.1)+
  geom_text(mapping=aes(x=1, y=0.6, label=pLabel), data=df_correlation, size=cortext_size)+
  ggtitle('SCORE DIFFERENCE')+
  theme_classic()+
  theme(plot.title = element_text(hjust=0.5,face='bold'))+
  xlab("")+
  ylab("Difference (Plausible - Implausible)")
plot.score_difference
savename <- "scoreDifference.png"
ggsave(paste(savedir,savename,sep="/"), width=20, height=6, units='cm')
```


#REGRESSION

## Stats Main

```{r regression, echo=FALSE}

metrics = levels(dat$Metric)
results = data.frame()

for(i in seq_along(metrics)){
  metric = metrics[i]
  print(metric)
  dat.metric = dat %>% filter(Metric==metric)
  m = lmer(FinalNormScore ~ Plausibility + agent_freq_norm + patient_freq_norm + verb_freq_norm + sentence_freq_norm + (1|ItemNum), 
           data=dat.metric, REML=FALSE, 
           control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e6)))
 d = data.frame(coef(summary(m)))
 d = d %>% rename(pVal=Pr...t..)
 d$Metric = metric 
 d$Parameter = rownames(d)
 d = d[,c(6,7,1:5)]
 results = rbind(results, d)
}
```

## Adjust for multiple comparisons

```{r adjust pVals, echo=FALSE}
results = results %>%
  mutate(Category = ifelse(Metric%in%anns, "ANNs", "baselines")) %>%
  mutate(Category = ifelse(Metric=="human", "human", Category))

results = results %>%
  group_by(Category, Parameter) %>%
  mutate(pValAdjusted = p.adjust(pVal, method="fdr", n=length(pVal))) %>%
  mutate(pLabel = ifelse(pValAdjusted<0.001, "***", 
                         ifelse(pValAdjusted<0.01, "**",
                                ifelse(pValAdjusted<0.05, "*", "")))) %>%
  mutate(pValFinal = ifelse(pValAdjusted<0.001, "<0.001", as.character(round(pValAdjusted,3)))) %>%
  ungroup()

savename <- "regression_by_metric.csv"
write.csv(results, paste(savedir,savename,sep="/"), row.names=FALSE)
```

