---
title: "results_main"
output: html_document
---

# SETUP

```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls()) 
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(grid)
library(gridExtra)
library(operator.tools)
library(lme4)
library(lmerTest)
library(cocor)
library(patchwork)
library(gtools)

source('dataloader_utils.R') #includes normalizations, read_data functions
source('stats_utils.R')

# Suppress summarise info
options(dplyr.summarise.inform = FALSE)
```

### Print environment variables 
```{r}
#SET ENVIRONMENT VARIABLES
which_models <- "all_main" #can be "anns_all", "anns_main", or "all_main"
# anns_main = human + main ANNs
# all_main = human + main ANNs + main baselines
# anns_all = human + all ANNs
normalization <- "min-max" #can be "min-max", "zscore"

path <- paste("results/")
ifelse(!dir.exists(path), dir.create(path), FALSE)
savedir <- paste(path,"EventsAdapt_Models=",which_models,sep='')
ifelse(!dir.exists(savedir), dir.create(savedir), FALSE)

print(paste(savedir))
message("Running with the following environment variables:")
message(paste("which_models: ", which_models))
message(paste("normalization: ", normalization))

#for plotting the dotted reference line
if (grepl("min-max", normalization) == TRUE){
  reference_value = 0.5
} else {
  reference_value = 0
}
```

# READ DATA
```{r}
#created via preprocess_datasets_Events*.Rmd
dat = read.csv('clean_data/clean_EventsAdapt_df.csv')
human_dat = read.csv('clean_data/clean_EventsAdapt_human_dat.csv')
```

```{r}
if (grepl("min-max", normalization) == TRUE){
  normalization <- function(x) {
    return(min_max(x)) }
} else {
  normalization <- function(x) {
    return(scale_this(x)) }
}
```

# Select which models/metrics to plot

```{r, utils.choose_models, echo=FALSE}
human <- c("human")
anns <- c("GPT2","BERT","RoBERTa","tinyLSTM")
anns_main <- c("GPT2-XL.l2r","BERT-large.pverb","RoBERTa-large.pverb","tinyLSTM.surprisal") #best metrics
baselines <- c("syntax.PPMI","SDM")

anns_main <- c(human, anns_main)
all_main <- c(anns_main, baselines)

unique_metrics = unique(dat$Metric)
anns <- unique_metrics[grepl(paste(anns, collapse = "|"), unique_metrics)]
anns = mixedsort(anns)

anns_all <- c(human,anns)

# select which to use
chosen_models = eval(parse(text=which_models))
dat = dat %>% filter(Metric %in% chosen_models)

message("Using these models/metrics:")
dat$Metric = factor(dat$Metric, levels = chosen_models)
print(unique(dat$Metric))
```

# CONTRASTS
```{r}
dat$Plausibility = factor(dat$Plausibility, levels=c("Plausible", "Implausible")) # dummy coding by default

dat$Voice = as.factor(dat$Voice)
contrasts(dat$Voice) = c(0.5, -0.5)
colnames(attr(dat$Voice, "contrasts")) = c("A>P")

dat$TrialType = factor(dat$TrialType, levels=c("AAR", "AAN", "AI"))    # dummy coding by default
dat = within(dat, Metric <- relevel(Metric, ref = "human"))    # set humans as the reference 
dat$ItemNum = as.factor(dat$ItemNum)
```


# BINARY ACCURACY

## General plotting settings

```{r}
#Set plotting options for grid plots
nr_models = length(chosen_models)
ncols=nr_models #models plus human
nrows=round(nr_models/ncols)
if (nrows * ncols < nr_models) {
  nrows = nrows + 1
}
```

```{r, echo=FALSE, fig.height=12, fig.width=15}

#add Category color
if (which_models == "anns_all") {
  dat.binchoice = dat %>%
  mutate(Metric = as.character(Metric)) %>%
  mutate(Category = ifelse(startsWith(Metric, "BERT"), "BERT", "RoBERTa")) %>%
  mutate(Category = ifelse(startsWith(Metric, "GPT2"), "GPT2", Category)) %>%
  mutate(Category = ifelse(startsWith(Metric, "tiny"), "tinyLSTM", Category)) %>%
  mutate(Category = ifelse(Metric=="human", "human", Category))
} else {
  dat.binchoice = dat %>%
  mutate(Category = ifelse(Metric%in%anns, "ANNs", "baselines")) %>%
  mutate(Category = ifelse(Metric=="human", "human", Category))
}

dat.binchoice = dat.binchoice %>%
  group_by(ItemNum, TrialType, Voice, Metric, LowerBetter, Category) %>%
  summarize(ScoreDiff = NormScore[Plausibility=="Plausible"]-NormScore[Plausibility=="Implausible"]) %>%
  mutate(FinalScoreDiff = ifelse(LowerBetter==TRUE, -ScoreDiff, ScoreDiff)) %>%
  mutate(Accuracy = ifelse(FinalScoreDiff>0, 1, 0)) %>%
  ungroup()

# specify order for plotting
dat.binchoice$Metric = factor(dat.binchoice$Metric, levels=chosen_models)
dat.binchoice$TrialType = factor(dat.binchoice$TrialType, levels=c("AI", "AAN", "AAR"))

if (which_models=="anns_all") {
  dat.binchoice$Category = factor(dat.binchoice$Category, levels=c("human", "GPT2", "BERT", "RoBERTa", "tinyLSTM"))
} else {
  dat.binchoice$Category = factor(dat.binchoice$Category, levels=c("human", "ANNs", "baselines"))
}
```

### Stats

```{r}
# get p values
dat.binchoice.summary = dat.binchoice %>%
  filter(TrialType %in% c("AI", "AAN")) %>%
  group_by(Category, Metric, TrialType) %>%
  summarize(NumCorrect=sum(Accuracy), NumTotal=length(Accuracy)) %>%
  mutate(AccuracyScore = NumCorrect/NumTotal) %>%
  ungroup() %>%
  mutate(pVal = calculate_binom_pval(NumCorrect, NumTotal))

# adjust for multiple comparisons within each category
dat.binchoice.summary = dat.binchoice.summary %>%
  group_by(Category) %>%
  mutate(pValAdjusted = p.adjust(pVal, method="fdr", n=length(pVal)),
         ntoadjust = length(pVal)) %>%
  mutate(pLabel = ifelse(pValAdjusted<0.001, "***", 
                         ifelse(pValAdjusted<0.01, "**",
                                ifelse(pValAdjusted<0.05, "*", "")))) %>%
  ungroup()
```

```{r}
# get human responses separately
human.results = dat.binchoice.summary %>%
  filter(Metric=='human') %>%
  select(TrialType, NumCorrect, NumTotal) %>%
  rename(NumCorrectHuman=NumCorrect, NumTotalHuman=NumTotal)
dat.binchoice.summary.withchisq = merge(dat.binchoice.summary, human.results)
dat.binchoice.summary.withchisq = dat.binchoice.summary.withchisq %>%
  mutate(ChiSq = calculate_chisq_vectorized_chi(NumCorrect, NumTotal, NumCorrectHuman),
         pVal2humans = calculate_chisq_vectorized_p(NumCorrect, NumTotal, NumCorrectHuman)) %>%
  group_by(Category) %>%
  mutate(pVal2humansAdjusted = p.adjust(pVal2humans, method="fdr", n=length(pVal2humans)),
         ntoadjust = length(pVal2humans)) %>%
  mutate(pLabel2humans = ifelse(pVal2humansAdjusted<0.001, "***", 
                         ifelse(pVal2humansAdjusted<0.01, "**",
                                ifelse(pVal2humansAdjusted<0.05, "*", "")))) 
```

### Plot

```{r, echo=FALSE}

plot.binacc = ggplot(data=subset(dat.binchoice, TrialType%in%c("AI","AAN")), 
       mapping=aes(x=Metric, y=Accuracy, fill=Category))+
  facet_wrap(~TrialType, ncol = 1)+
  geom_hline(yintercept=1, color='gray50', linetype='dotted')+
  stat_summary(geom='col', fun.y='mean',
               color='black', width=0.8)+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', size = 0.5, width=0.1)+
  geom_text(mapping=aes(x=Metric, y=0.05, label=pLabel), data=dat.binchoice.summary)+
  coord_cartesian(ylim=c(0,1))+
  geom_hline(yintercept=.5, linetype='dotted')+
  theme_classic()+
  labs(x='Model')+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
plot.binacc

savename <- "1_binaryAccuracy_split.png"
ggsave(paste(savedir,savename,sep="/"), width=15, height=12, units='cm')
```
### Stats: difference between AI & AAN conditions

Hypothesis test of proportions (Two proportions z test):
prop.test(c(nr_successes1, nr_successes2), c(nr_total1, nr_total2), p = NULL, alternative = "two.sided",
          correct = TRUE)
          
The function returns:
- the value of Pearson’s chi-squared test statistic.
- a p-value
- a 95% confidence intervals
- an estimated probability of success (the proportion of smokers in the two groups)

```{r, echo=FALSE}
dat.aivsaan = dat.binchoice.summary %>% 
  dplyr::select(TrialType, Metric, NumCorrect, NumTotal)

df_ai.vs.aan = data.frame()

message(paste("Hypothesis test for proportions | AI vs. AAN accuracy"))
for (i in seq_along(chosen_models)) {
  curr_data = subset(dat.aivsaan, Metric==chosen_models[i])
  cnt_successes = curr_data$NumCorrect
  cnt_total = curr_data$NumTotal
  
  proptest <- prop.test(x = cnt_successes, n = cnt_total)
  
  proptest.stat = proptest$statistic
  proptest.pval = proptest$p.value
  
  print(paste(chosen_models[i], "|", proptest.stat, "|", proptest.pval, "|", plabel(proptest.pval)))
  
  df <- data.frame(chosen_models[i], proptest.stat, proptest.pval, plabel(proptest.pval))
  df_ai.vs.aan <- rbind(df_ai.vs.aan,df)
}

colnames(df_ai.vs.aan) = c("Metric", "statistic", "pVal", "pLabel")
```

### Extra: Plot AAR binary accuracy
```{r, echo=FALSE}

# get p values
dat.binchoice.summary.aar = dat.binchoice %>%
  filter(TrialType %in% c("AAR")) %>%
  group_by(Category, Metric, TrialType) %>%
  summarize(NumCorrect=sum(Accuracy), NumTotal=length(Accuracy)) %>%
  ungroup() %>%
  mutate(pVal = calculate_binom_pval(NumCorrect, NumTotal))

# adjust for multiple comparisons within each category
dat.binchoice.summary.aar = dat.binchoice.summary.aar %>%
  group_by(Category) %>%
  mutate(pValAdjusted = p.adjust(pVal, method="fdr", n=length(pVal)),
         ntoadjust = length(pVal)) %>%
  mutate(pLabel = ifelse(pValAdjusted<0.001, "***", 
                         ifelse(pValAdjusted<0.01, "**",
                                ifelse(pValAdjusted<0.05, "*", ""))))

#check for AAR condition (should be around chance)
plot.binchoice.aar =ggplot(data=subset(dat.binchoice, TrialType%in%c("AAR")), 
       mapping=aes(x=Metric, y=Accuracy, fill=Category))+
  facet_wrap(~TrialType, ncol = 1)+
  geom_hline(yintercept=1, color='gray50', linetype='dotted')+
  stat_summary(geom='col', fun.y='mean',
               color='black', width=0.8)+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', size = 0.5, width=0.1)+
  geom_text(mapping=aes(x=Metric, y=0.05, label=pLabel), data=dat.binchoice.summary.aar)+
  coord_cartesian(ylim=c(0,1))+
  geom_hline(yintercept=.5, linetype='dotted')+
  theme_classic()+
  labs(x='Model')+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
plot.binchoice.aar

savename <- "1_binaryAccuracy_split_AAR.png"
ggsave(paste(savedir,savename,sep="/"), width=15, height=12, units='cm')
```

## Error analysis

```{r}
acc_by_sentence.ANN = dat.binchoice %>% 
  filter(TrialType %in% c("AI", "AAN")) %>%
  filter(Category=='ANNs') %>%
  group_by(ItemNum, Voice) %>%
  mutate(meanAccuracy=mean(Accuracy))
```

### Quick plots
```{r}
ggplot(acc_by_sentence.ANN, aes(x=meanAccuracy)) + geom_histogram() + 
  facet_wrap(~Voice)
```
```{r}
# can adjust bin width slightly to show it's actually evenly distributed
ggplot(acc_by_sentence.ANN, aes(x=meanAccuracy, fill=Voice)) + geom_histogram(binwidth=0.1, position="dodge") 
```

### Save hard sentences

```{r}
hard_sents = acc_by_sentence.ANN %>%
  filter(meanAccuracy<0.5) %>%
  select(ItemNum, Voice, meanAccuracy)

hard_sents = inner_join(hard_sents, dat) %>%
  select(ItemNum, Sentence, Voice, Plausibility, meanAccuracy) %>% 
  distinct() %>%
  spread(Plausibility, Sentence)

savename <- "hard_sentences_ANNs.csv"
write.csv(hard_sents, paste(savedir,savename,sep="/"), row.names=FALSE)
```

### Compare with human scores

```{r}
dat.binchoice.human = dat.binchoice %>% filter(Metric=='human')
dat.binchoice.human = inner_join(dat.binchoice.human, acc_by_sentence.ANN %>% select(ItemNum, TrialType, Voice, meanAccuracy)) %>%
  distinct()

ggplot(dat.binchoice.human, aes(x=meanAccuracy, y=FinalScoreDiff)) + 
  stat_summary(geom='col', fun.y='mean')+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', size = 0.5, width=0.1)
```

TODO: regression stats

```{r}
ggplot(dat.binchoice.human, aes(x=meanAccuracy, y=FinalScoreDiff)) + 
  geom_jitter(width=0.02)

cor(dat.binchoice.human$meanAccuracy, dat.binchoice.human$FinalScoreDiff)
```

# MODEL HUMAN COMPARISON

## Calculate human 2 human response correlation 
(each human with the average of all the rest)
```{r}
human_dat.comp = human_dat %>%
  mutate(FinalNormScore = normalization(Score)) %>%
  select(WorkerId, ItemNum, TrialType, Plausibility, Voice, Sentence, FinalNormScore)

workerIDs = unique(human_dat.comp$WorkerId)
df_correlation = data.frame()

for (i in seq_along(workerIDs)) {
  if (i%%50==0) {    # just to keep track (n=511 for EventsAdapt)
    print(i)
  }
  human_dat.this = human_dat.comp %>% filter(WorkerId==workerIDs[i])
  human_dat.rest = human_dat.comp %>% filter(WorkerId!=workerIDs[i]) %>%
    filter(ItemNum %in% unique(human_dat.this$ItemNum)) %>%
    group_by(ItemNum, Plausibility, Voice, Sentence) %>% 
    summarize(MeanNormScore=mean(FinalNormScore))
  
  human_dat.side2side = merge(human_dat.this, human_dat.rest,
                              by=c("ItemNum", "Plausibility", "Voice"))
  corval= cor(human_dat.side2side$FinalNormScore, human_dat.side2side$MeanNormScore)
  # add vector to a dataframe
  df <- data.frame(workerIDs[i], corval)
  df_correlation <- rbind(df_correlation,df)
}

human2human_corr = mean(df_correlation$corval)
print(paste("The human2human overall correlation value is", human2human_corr))
```

## Sample a random human response for each item 
And compute the avg of all other responses 

```{r}
set.seed(21)
human_dat.sampled = human_dat.comp %>%
  group_by(ItemNum, TrialType, Plausibility, Voice, Sentence) %>%
  summarize(RandomScore = sample(FinalNormScore,1),
            MeanScoreRest = (sum(FinalNormScore)-RandomScore)/(length(FinalNormScore)-1)) %>%
  ungroup()
```

## Prep data

```{r}
dat.human = dat %>% filter(Metric=="human") %>%
  select(ItemNum, TrialType, Sentence, Plausibility, FinalNormScore)
dat.model = dat %>% filter(Metric!="human") %>%
  select(ItemNum, TrialType, Sentence, Plausibility, FinalNormScore, Metric)
dat.model2human = merge(dat.human, dat.model, 
                        by=c("ItemNum", "TrialType", "Sentence", "Plausibility")) %>%
  rename(HumanScore=FinalNormScore.x, ModelScore=FinalNormScore.y)
```

```{r}
# add human2human data
human_dat.sampled = human_dat.sampled %>%
  select(-Voice) %>%
  filter(Sentence %in% unique(dat.model2human$Sentence)) %>%  # make sure it's the same sents
  rename(HumanScore=MeanScoreRest, ModelScore=RandomScore)    # renaming for convenience 
human_dat.sampled$Metric = 'singleHuman'
```

```{r}
dat.model2human = rbind(dat.model2human, human_dat.sampled)
```

Formatting:

```{r, echo=FALSE, fig.height=12, fig.width=15}
dat.model2human = dat.model2human %>%
  #add color info
  mutate(Category = ifelse(Metric%in%anns, "ANNs", "baselines")) %>%
  mutate(Category = ifelse(Metric=="singleHuman", "human", Category))

# specify order for plotting
metric_names = chosen_models
metric_names[1] = "singleHuman"
dat.model2human$Metric = factor(dat.model2human$Metric, levels=metric_names)
dat.model2human$Category = factor(dat.model2human$Category, levels=c("human", "ANNs", "baselines"))
dat.model2human$TrialType = factor(dat.model2human$TrialType, levels=c("AI", "AAN", "AAR"))
```

## Calculate correlations for the plot

```{r}
trialTypes = c("AI", "AAN", "AAR")
df_correlation = get_correlation_df_tt("model2human", "singleHuman", trialTypes, dat.model2human, metric_names)
```

## Plot

```{r}

dat.model2human$TrialType = factor(dat.model2human$TrialType, levels=c("AI","AAN","AAR"))
df_correlation$Metric = factor(df_correlation$Metric, levels=metric_names)
df_correlation$TrialType = factor(df_correlation$TrialType, levels=c("AI","AAN","AAR"))
df_correlation$Metric = factor(df_correlation$Metric, levels=metric_names)

# text size 
cortext_size = 3

plot.model2human = ggplot(dat=dat.model2human)+  
  geom_point(mapping=aes(x=ModelScore, y=HumanScore, color=Category), 
             size=0.1, position=position_jitter(width=0.01))+
  geom_abline(slope=1, intercept=0, size=0.2)+
  geom_point(mapping=aes(x=ModelScore, y=HumanScore, color=Category), 
             size=0.1, position=position_jitter(width=0.01))+
  geom_text(mapping=aes(x=0.2, y=1.10, 
                        label=sprintf("r = %.2f%s", round(Correlation,2), pVal2zeroLabel)), 
            data=df_correlation, size=cortext_size, hjust = 0)+
  coord_cartesian(ylim=c(0,1.15), xlim=c(-0.15,1))+
  scale_x_continuous(breaks=c(0,0.5,1))+
  scale_y_continuous(breaks=c(0,0.5,1))+
  theme_classic()+
  facet_grid(TrialType~Metric)
plot.model2human

savename <- "model2human_byTrialType_ANNs.png"
ggsave(paste(savedir,savename,sep="/"), height=12,width=24, units='cm')
```

```{r}

dat.model2human.sub = dat.model2human %>% 
  filter(TrialType %in% c("AI","AAN"))

df_correlation = df_correlation %>%
  filter(TrialType %in% c("AI","AAN"))

dat.model2human.sub$TrialType = factor(dat.model2human.sub$TrialType, levels=c("AI","AAN","AAR"))
df_correlation$Metric = factor(df_correlation$Metric, levels=metric_names)
df_correlation$TrialType = factor(df_correlation$TrialType, levels=c("AI","AAN","AAR"))
df_correlation$Metric = factor(df_correlation$Metric, levels=metric_names)

# text size 
cortext_size = 3

plot.model2human.sub = ggplot(dat=dat.model2human.sub)+  
  geom_point(mapping=aes(x=ModelScore, y=HumanScore, color=Category), 
             size=0.1, position=position_jitter(width=0.01))+
  geom_abline(slope=1, intercept=0, size=0.2)+
  geom_point(mapping=aes(x=ModelScore, y=HumanScore, color=Category), 
             size=0.1, position=position_jitter(width=0.01))+
  geom_text(mapping=aes(x=-0.25, y=1, 
                        label=sprintf("r = %.2f%s", round(Correlation,2), pVal2zeroLabel)), 
            data=df_correlation, size=cortext_size, hjust = 0)+
  coord_cartesian(ylim=c(0,1), xlim=c(-0.2,1))+
  theme_classic()+
  facet_grid(TrialType~Metric)
plot.model2human.sub

savename <- "model2human_byTrialType_ANNs_noAAR.png"
ggsave(paste(savedir,savename,sep="/"), height=10,width=25, units='cm')
```


# SCATTERPLOTS

## General settings
```{r}

# color scheme
color_plaus = '#1b9e77'
color_voice = '#d95f02'
color_syn = '#7570b3'

```

## PLAUSIBLE VS IMPLAUSIBLE

### Prep data
```{r, echo=FALSE}
dat.plaus = dat %>% 
  dplyr::select(ItemNum, TrialType, Plausibility, Voice, Metric, FinalNormScore) %>%
  group_by(ItemNum, TrialType, Plausibility, Voice, Metric) %>%
  summarize(meanScore = mean(FinalNormScore, na.rm=TRUE)) %>%
  spread(Plausibility, meanScore)  
```

### Calculate correlations


```{r}
print(paste(chosen_models))
``` 

```{r}
trialTypes = c('AI', 'AAN')
dat.plaus.sub = dat.plaus %>%
  filter(TrialType%in%c(trialTypes))
df_correlation.plaus = get_correlation_df("plausibility", "human", dat.plaus.sub, chosen_models)
```

### Plot
```{r}
#determine plotting order
df_correlation.plaus$Metric = factor(df_correlation.plaus$Metric, levels=chosen_models)
dat.plaus.sub$Metric = factor(dat.plaus.sub$Metric, levels=chosen_models)

plot.plaus = ggplot(data=dat.plaus.sub)+
  geom_point(mapping=aes(x=Plausible, y=Implausible), size=0.2, color=color_plaus)+
  #facet_wrap(~Metric, ncol=ncols, nrow=nrows)+
  geom_abline(slope=1, intercept=0, size=0.2)+
  geom_point(mapping=aes(x=Plausible, y=Implausible), size=0.2, color=color_plaus)+
  geom_text(mapping=aes(x=0, y=1.10, 
                        label=sprintf("r = %.2f%s", round(Correlation,2), pVal2zeroLabel)), 
            data=df_correlation.plaus, size=cortext_size, hjust = 0)+
  coord_cartesian(ylim=c(0,1.15), xlim=c(0,1))+
  scale_x_continuous(breaks=c(0,0.5,1))+
  scale_y_continuous(breaks=c(0,0.5,1))+
  ggtitle('PLAUSIBILITY')+
  theme_classic()+
  facet_grid(~Metric)+
  theme(plot.title = element_text(hjust=0.5,face='bold'))
plot.plaus
#paste("r",round(Correlation,2),sep=" = ")
```

```{r, fig.height=2,fig.width=5}
dat.sub = dat %>%
  filter(TrialType%in%c("AI","AAN"))

df_correlation.plaus$Metric = factor(df_correlation.plaus$Metric, levels=chosen_models)
dat.sub$Metric = factor(dat.sub$Metric, levels=chosen_models)

plot.plaus.density.sub = ggplot()+
  geom_density(data=dat.sub, mapping=aes(FinalNormScore, fill = factor(Plausibility, levels=c("Implausible","Plausible"))), alpha = 0.2)+
  facet_wrap(~Metric,nrow=nrows, ncol=ncols)+
  geom_text(mapping=aes(x=0.1, y=5, label=sprintf("r = %.2f%s", round(Correlation,2), pVal2zeroLabel)),  data=df_correlation.plaus, size=cortext_size, hjust = 0)+
  ggtitle('PLAUSIBILITY')+
  geom_vline(xintercept=reference_value, linetype='dotted')+
  #coord_flip()+
  theme_classic()+
  xlab("Normed score") +
  ylab("Density")+
  theme(plot.title = element_text(hjust=0.5,face='bold'))+
  theme(legend.title = element_blank())
plot.plaus.density.sub
```
#AAR density
```{r}
df_correlation.plaus = data.frame()

dat.plaus.sub = dat.plaus %>%
  filter(TrialType%in%c("AAR")) #subsetting here for plausibility as 
df_correlation.plaus = get_correlation_df("plausibility", "human", dat.plaus.sub, chosen_models)
```


```{r, fig.height=2,fig.width=5}
dat.sub = dat %>%
  filter(TrialType%in%c("AAR"))

df_correlation.plaus$Metric = factor(df_correlation.plaus$Metric, levels=chosen_models)
dat.sub$Metric = factor(dat.sub$Metric, levels=chosen_models)

plot.plaus.density.sub.aar = ggplot()+
  geom_density(data=dat.sub, mapping=aes(FinalNormScore, fill = Plausibility), alpha = 0.2)+
  facet_wrap(~Metric,nrow=nrows, ncol=ncols)+
  geom_text(mapping=aes(x=0.1, y=5, label=sprintf("r = %.2f%s", round(Correlation,2), pVal2zeroLabel)),  data=df_correlation.plaus, size=cortext_size, hjust = 0)+
  ggtitle('PLAUSIBILITY (BOTH VERSIONS PLAUSIBLE)')+
  geom_vline(xintercept=reference_value, linetype='dotted')+
  #coord_flip()+
  theme_classic()+
  xlab("Normed score") +
  ylab("Density")+
  theme(plot.title = element_text(hjust=0.5,face='bold'))+
  theme(legend.title = element_blank())+
  scale_fill_discrete(labels = c("Plausible (V1)", "Plausible (V2)"))
plot.plaus.density.sub.aar
```

```{r}
#NOTE: Here we're not necessarily interested in the correlation actually but rather if it's significantly above the reference value
#AAR control

df_correlation.plaus$Metric = factor(df_correlation.plaus$Metric, levels=chosen_models)
dat.plaus.sub$Metric = factor(dat.plaus.sub$Metric, levels=chosen_models)

plot.plaus.control = ggplot(data=dat.plaus.sub)+
  geom_point(mapping=aes(x=Plausible, y=Implausible), size=0.2, color=color_plaus)+
  facet_wrap(~Metric, ncol=ncols, nrow=nrows)+
  geom_hline(yintercept=reference_value, linetype='dotted')+
  geom_vline(xintercept=reference_value, linetype='dotted')+
  geom_abline(slope=1, intercept=0, size=0.2)+
  geom_point(mapping=aes(x=Plausible, y=Implausible), size=0.2, color=color_plaus)+
  geom_point(mapping=aes(x=Plausible, y=Implausible), size=0.2, color=color_plaus)+
  geom_text(mapping=aes(x=0, y=1, label=sprintf("r = %.2f%s", round(Correlation,2), pVal2zeroLabel)), 
            data=df_correlation.plaus, size=cortext_size, hjust = 0)+
  coord_cartesian(ylim=c(0,1), xlim=c(0,1))+
  ggtitle('PLAUSIBILITY (BOTH VERSIONS PLAUSIBLE)')+
  theme_classic()+
  xlab("Plausible V1") +
  ylab("Plausible V2") +
  theme(plot.title = element_text(hjust=0.5,face='bold'))
plot.plaus.control

savename <- "scatterplot_plausibility_aar.png"
ggsave(paste(savedir,savename,sep="/"), height=12,width=28, units='cm')
#paste("r",round(Correlation,2),sep=" = ")
```
#Get ratio (how many above mean rating) AAR & STATS
```{r}
dat.plaus.aar = dat %>%
  filter(Metric%in%chosen_models) %>%
  filter(TrialType=="AAR") %>%
  mutate(Greater05 = ifelse(FinalNormScore > 0.5,1,0)) %>%
  group_by(Metric) %>%
  summarize(AARCorrect = sum(Greater05)/length(Greater05))

dat.plaus.aar.stats = dat %>%
  filter(Metric%in%chosen_models) %>%
  filter(TrialType=="AAR") %>%
  mutate(AdjustedScore = FinalNormScore-0.5)

df_ttest = data.frame()
for(i in seq_along(chosen_models)){
  curr_data = subset(dat.plaus.aar.stats, Metric==chosen_models[i])
  tt = t.test(curr_data$AdjustedScore)
  pval2zero = tt$p.value
  ttstats = tt$statistic
  df <- data.frame(chosen_models[i], ttstats, pval2zero)
    df_ttest <- rbind(df_ttest,df)
}
```



```{r, echo=FALSE}
dat.binchoice.aar = subset(dat.binchoice, TrialType%in%c("AAR"))
dat.binchoice.aar = subset(dat.binchoice.aar, Metric %in% chosen_models)
dat.binchoice.aar$Metric = factor(dat.binchoice.aar$Metric, levels=chosen_models)

df_correlation = data.frame()
# get p values for t test
for(i in seq_along(chosen_models)){
  curr_dat = subset(dat.binchoice.aar,Metric==chosen_models[i])
  curr_score_diff = c(curr_dat$FinalScoreDiff)
  print(paste(t.test(curr_score_diff)$p.value))
  ttestPval2zero = t.test(curr_score_diff)$p.value
  print(paste(chosen_models[i], ttestPval2zero))
  curr_cat = unique(curr_dat$Category)
  # add vector to a dataframe
    df <- data.frame(chosen_models[i], ttestPval2zero,curr_cat)
    df_correlation <- rbind(df_correlation,df)
}

# adjust for multiple comparisons within each category
df_correlation = df_correlation %>%
  mutate(pLabel = ifelse(ttestPval2zero<0.001, "***", 
                         ifelse(ttestPval2zero<0.01, "**",
                                ifelse(ttestPval2zero<0.05, "*",
                                       ifelse(ttestPval2zero>=0.05, "n.s.", "")))))

colnames(df_correlation) = c("Metric", "ttestPval2zero", "Category", "pLabel")

df_correlation$Metric = factor(df_correlation$Metric, levels=chosen_models)
dat.binchoice.aar$Metric = factor(dat.binchoice.aar$Metric, levels=chosen_models)

plot.score_difference.aar = ggplot(data=dat.binchoice.aar,mapping=aes(x=TrialType, y=FinalScoreDiff, fill=Category))+
  facet_wrap(~Metric, ncol=ncols, nrow=nrows)+
  stat_summary(geom='col', fun.y='mean',
               color='black', width=0.5)+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', size = 0.5, width=0.1)+
  geom_text(mapping=aes(x=1, y=0.015, label=pLabel), data=df_correlation, size=cortext_size)+
  ggtitle('SCORE DIFFERENCE')+
  theme_classic()+
  theme(plot.title = element_text(hjust=0.5,face='bold'))+
  xlab("")+
  ylab("Difference (V1 - V2)")+
  coord_cartesian(ylim=c(-0.01,0.015))
plot.score_difference.aar
savename <- "scoreDifference_AAR.png"
ggsave(paste(savedir,savename,sep="/"), width=20, height=6, units='cm')
```


## COMBINE: Control plot AAR

```{r}
plot.plaus.control / plot.score_difference.aar 

savename <- "controlplot_aar.png"
ggsave(paste(savedir,savename,sep="/"), height=14,width=28, units='cm')
```


## ACTIVE VS PASSIVE

### Prep data

```{r, echo=FALSE}
dat.voice = dat %>%
  filter(Experiment=='EventsAdapt') %>%
  select(ItemNum, TrialType, Plausibility, Voice, Metric, FinalNormScore) %>%
  group_by(ItemNum, TrialType, Plausibility, Voice, Metric) %>%
  summarize(meanScore = mean(FinalNormScore, na.rm=TRUE)) %>%
  spread(Voice, meanScore)
dat.voice$Metric = factor(dat.voice$Metric, levels=chosen_models)
```

### Calculate correlations

```{r}
df_correlation.voice = get_correlation_df("voice", "human", dat.voice, chosen_models)
```

### Plot

```{r}
df_correlation.voice$Metric = factor(df_correlation.voice$Metric, levels=chosen_models)
dat.voice$Metric = factor(dat.voice$Metric, levels=chosen_models)

plot.voice = ggplot(data=dat.voice)+
  geom_point(mapping=aes(x=active, y=passive), size=0.2, color=color_voice)+
  facet_wrap(~Metric, ncol=ncols, nrow=nrows)+
  geom_abline(slope=1, intercept=0, size=0.2)+
  geom_text(mapping=aes(x=0, y=1.10, 
                        label=sprintf("r = %.2f%s", round(Correlation,2), pVal2zeroLabel)), 
            data=df_correlation.voice, size=cortext_size, hjust = 0)+
  coord_cartesian(ylim=c(0,1.15), xlim=c(0,1))+
  scale_x_continuous(breaks=c(0,0.5,1))+
  scale_y_continuous(breaks=c(0,0.5,1))+
  theme_classic()+
  facet_grid(~Metric)+
  ggtitle('VOICE')+
  xlab("Active")+
  ylab("Passive")+
  theme(plot.title = element_text(hjust=0.5,face='bold'))
plot.voice
```


## SYNONYMOUS SENTENCES

### Prep data

```{r, echo=FALSE}
dat.syn = dat %>%
  filter(!is.na(SynonymPair)) %>%
  dplyr::select(-ItemNum) %>%
  group_by(TrialType, NumSyn, SynonymPair, Plausibility, Voice, Metric) %>%
  summarize(meanScore = mean(FinalNormScore, na.rm=TRUE)) %>%
  spread(NumSyn, meanScore)

# remove the sentences that don't have synonyms
dat.syn = dat.syn %>%
  filter(!is.na(Version1)) %>%
  filter(!is.na(Version2))

dat.syn$Metric = factor(dat.syn$Metric, levels=chosen_models)
```

### Calculate correlations

```{r}
df_correlation.syn = get_correlation_df("synonym", "human", dat.syn, chosen_models)
```

### Plot


```{r}
df_correlation.syn$Metric = factor(df_correlation.syn$Metric, levels=chosen_models)
dat.syn$Metric = factor(dat.syn$Metric, levels=chosen_models)

plot.syn = ggplot(data=dat.syn)+
  geom_point(mapping=aes(x=Version1, y=Version2), size=0.2, color=color_syn)+
  #facet_wrap(~Metric, ncol=ncols, nrow=nrows)+
  geom_abline(slope=1, intercept=0, size=0.2)+
  geom_point(mapping=aes(x=Version1, y=Version2), size=0.2, color=color_syn)+
  geom_text(mapping=aes(x=0, y=1.10, 
                        label=sprintf("r = %.2f%s", round(Correlation,2), pVal2zeroLabel)), 
            data=df_correlation.syn, size=cortext_size, hjust = 0)+
  coord_cartesian(ylim=c(0,1.15), xlim=c(0,1))+
  scale_x_continuous(breaks=c(0,0.5,1))+
  scale_y_continuous(breaks=c(0,0.5,1))+
  theme_classic()+
  facet_grid(~Metric)+
  ggtitle('SYNONYMOUS SENTENCES')+
  theme(plot.title = element_text(hjust=0.5,face='bold'))
plot.syn
```


## COMBINE

```{r}
plot.plaus / plot.plaus.control 

savename <- "scatterplot_plausibility.png"
ggsave(paste(savedir,savename,sep="/"), height=14,width=28, units='cm')
```
```{r}
plot.plaus.control / plot.score_difference.aar 

savename <- "controlplot_aar.png"
ggsave(paste(savedir,savename,sep="/"), height=14,width=28, units='cm')
```

```{r}
plot.plaus.density.sub / plot.plaus.density.sub.aar / plot.voice / plot.syn

savename <- "scatterplot_density.png"
ggsave(paste(savedir,savename,sep="/"), height=24,width=28, units='cm')
```


```{r}
plot.plaus / plot.voice / plot.syn 

savename <- "scatterplot_all_labeled.png"
ggsave(paste(savedir,savename,sep="/"), height=24,width=28, units='cm')
```

```{r}
plot.plaus / plot.plaus.control / plot.voice / plot.syn 

savename <- "scatterplot_all_labeled_with_aar.png"
ggsave(paste(savedir,savename,sep="/"), height=32,width=28, units='cm')
```

# ADDITIONAL PLOTS

## Word frequency (sentence)

```{r}
plot.freq = ggplot(data=dat)+
  geom_point(mapping=aes(x=sentence_freq_norm, y=FinalNormScore), size=0.2)+
  facet_wrap(~Metric, ncol=ncols, nrow=nrows)+
  geom_abline(slope=1, intercept=0, size=0.2)+
  coord_cartesian(ylim=c(0,1), xlim=c(0,1))+
  ggtitle('WORD FREQUENCY (SENTENCE)')+
  theme_classic()+
  theme(plot.title = element_text(hjust=0.5,face='bold'))
plot.freq
```

#REGRESSION

## Determine whether to include Plausibility by ItemNum random slope
```{r regression, echo=FALSE}
metrics = unique(dat$Metric)
pvals_likelihoodratiotest = c()

for(i in seq_along(metrics)){
  metric = metrics[i]
  dat.metric = dat %>% filter(Metric==metric)
  m1 = lmer(NormScore ~ Plausibility*TrialType*Voice + agent_freq_norm + patient_freq_norm + verb_freq_norm + sentence_freq_norm + (1+Plausibility|ItemNum), 
           data=dat.metric, REML=FALSE, 
           control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e6)))
  m0 = lmer(NormScore ~ Plausibility*TrialType*Voice + agent_freq_norm + patient_freq_norm + verb_freq_norm + sentence_freq_norm + (1|ItemNum), 
           data=dat.metric, REML=FALSE, 
           control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e6)))
  pvals_likelihoodratiotest = c(pvals_likelihoodratiotest, anova(m1, m0)[2,8])
}
print(pvals_likelihoodratiotest)
print(lapply(pvals_likelihoodratiotest, function(x) {x<0.05/length(pvals_likelihoodratiotest)}))    # apply Bonferroni correction
```

Conclusion: the model with the Plausibility x ItemNum slope is a better fit for most metrics, so will stick to it.

## Stats main

```{r regression, echo=FALSE}
metrics = levels(dat$Metric)
results = data.frame()

for(i in seq_along(metrics)){
  metric = metrics[i]
  dat.metric = dat %>% filter(Metric==metric)
  m = lmer(NormScore ~ Plausibility*TrialType*Voice + agent_freq_norm + patient_freq_norm + verb_freq_norm + sentence_freq_norm + (1+Plausibility|ItemNum), 
           data=dat.metric, REML=FALSE, 
           control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e6)))
 d = data.frame(coef(summary(m)))
 d = d %>% rename(pVal=Pr...t..)
 d$Metric = metric 
 d$Parameter = rownames(d)
 d = d[,c(6,7,1:5)]
 results = rbind(results, d)
}
```

## Correct for MC within ANNs and baselines

```{r adjust pVals, echo=FALSE}
results = results %>%
  mutate(Category = ifelse(Metric%in%anns, "ANNs", "baselines")) %>%
  mutate(Category = ifelse(Metric=="human", "human", Category))

results = results %>%
  group_by(Category, Parameter) %>%
  mutate(pValAdjusted = p.adjust(pVal, method="fdr", n=length(pVal))) %>%
  mutate(pLabel = ifelse(pValAdjusted<0.001, "***", 
                         ifelse(pValAdjusted<0.01, "**",
                                ifelse(pValAdjusted<0.05, "*", "")))) %>%
  mutate(pValFinal = ifelse(pValAdjusted<0.001, "<0.001", as.character(round(pValAdjusted,3)))) %>%
  ungroup()

savename <- "regression_by_metric.csv"
write.csv(results, paste(savedir,savename,sep="/"), row.names=FALSE)
```







