---
title: "results_main"
output: html_document
---

# SETUP

```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls()) 
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(grid)
library(gridExtra)
library(operator.tools)
library(lme4)
library(lmerTest)
library(cocor)
library(patchwork)
library(gtools)

source('dataloader_utils.R') #includes normalizations, read_data functions
source('stats_utils.R')

# Suppress summarise info
options(dplyr.summarise.inform = FALSE)
```

### Print environment variables 
```{r}
#SET ENVIRONMENT VARIABLES

#can be:
# llms_main = human + main LLMs
# all_main = human + main LLMs + baselines
# llms_all = human + all LLMs
which_models <- "all_main" 
normalization <- "min-max" #can be "min-max", "zscore"

path <- paste("results/")
ifelse(!dir.exists(path), dir.create(path), FALSE)
savedir <- paste(path,"EventsAdapt_Models=",which_models,sep='')
ifelse(!dir.exists(savedir), dir.create(savedir), FALSE)

print(paste(savedir))
message("Running with the following environment variables:")
message(paste("which_models: ", which_models))
message(paste("normalization: ", normalization))

#for plotting the dotted reference line
if (grepl("min-max", normalization) == TRUE){
  reference_value = 0.5
} else {
  reference_value = 0
}
```

# READ DATA
```{r}
#created via preprocess_datasets_Events*.Rmd
dat = read.csv('clean_data/clean_EventsAdapt_df.csv')
human_dat = read.csv('clean_data/clean_EventsAdapt_human_dat.csv')
```

```{r}
if (grepl("min-max", normalization) == TRUE){
  normalization <- function(x) { return(min_max(x)) }
} else {
  normalization <- function(x) { return(scale_this(x)) }
}
```

# Select which models/metrics to plot

```{r, utils.choose_models, echo=FALSE}
human <- c("human")
llms <- c("GPT2","BERT","RoBERTa")
llms_main <- c("GPT2-xl.l2r","BERT-large.pverb","RoBERTa-large.pverb")
baselines <- c("tinyLSTM.surprisal","syntax.PPMI","thematicFit.prod", "SDM")

llms_main <- c(human, llms_main)
all_main <- c(llms_main, baselines)

unique_metrics = unique(dat$Metric)
llms <- unique_metrics[grepl(paste(llms, collapse = "|"), unique_metrics)]
llms = mixedsort(llms)

llms_all <- c(human,llms)

# select which to use
chosen_models = eval(parse(text=which_models))
dat = dat %>% filter(Metric %in% chosen_models)

message("Using these models/metrics:")
dat$Metric = factor(dat$Metric, levels = chosen_models)
print(unique(dat$Metric))

# shorten model names
llms_main_short <- c("GPT2-xl","BERT-large","RoBERTa-large")
baselines_short <- c("tinyLSTM","syntax-PPMI","thematicFit", "SDM")
models_short = c(human, llms_main_short, baselines_short)
dat = dat %>%
  separate(Metric, c("Model", "ProbEstimate"), sep="\\.", remove=FALSE) %>%
  mutate(Model = replace(Model, Model=="syntax", "syntax-PPMI"))
dat$Model = factor(dat$Model, levels=models_short)
```

## Avg word frequency

```{r}
dat.sentfreq = dat %>% 
  filter(Voice=='active') %>%
  select(TrialType, Sentence, sentence_freq) %>%
  distinct() %>%
  group_by(TrialType) %>%
  summarize(avg_sent_freq = mean(sentence_freq))
```

# CONTRASTS
```{r}
dat$Plausibility = factor(dat$Plausibility, levels=c("Plausible", "Implausible")) # dummy coding by default

dat$Voice = as.factor(dat$Voice)
contrasts(dat$Voice) = c(0.5, -0.5)
colnames(attr(dat$Voice, "contrasts")) = c("A>P")

dat$TrialType = factor(dat$TrialType, levels=c("AAN", "AAR", "AI"))    # dummy coding by default
dat = within(dat, Metric <- relevel(Metric, ref = "human"))    # set humans as the reference 
dat$ItemNum = as.factor(dat$ItemNum)
```


# BINARY ACCURACY

## General plotting settings

```{r}
#Set plotting options for grid plots
nr_models = length(chosen_models)
ncols=nr_models #models plus human
nrows=round(nr_models/ncols)
if (nrows * ncols < nr_models) {
  nrows = nrows + 1
}

cortext_size = 4
```

```{r, echo=FALSE, fig.height=12, fig.width=15}

#add Category color
if (which_models == "llms_all") {
  dat.binchoice = dat %>%
  mutate(Metric = as.character(Metric)) %>%
  mutate(Category = ifelse(startsWith(Metric, "BERT"), "BERT", "RoBERTa")) %>%
  mutate(Category = ifelse(startsWith(Metric, "GPT2"), "GPT2", Category)) %>%
  mutate(Category = ifelse(Metric=="human", "human", Category))
} else {
  dat.binchoice = dat %>%
  mutate(Category = ifelse(Metric%in%llms, "LLMs", "baselines")) %>%
  mutate(Category = ifelse(Metric=="human", "human", Category))
}

dat.binchoice = dat.binchoice %>%
  group_by(ItemNum, TrialType, Voice, Metric, Model, LowerBetter, Category) %>%
  summarize(ScoreDiff = NormScore[Plausibility=="Plausible"]-NormScore[Plausibility=="Implausible"]) %>%
  mutate(FinalScoreDiff = ifelse(LowerBetter==TRUE, -ScoreDiff, ScoreDiff)) %>%
  mutate(Accuracy = ifelse(FinalScoreDiff>0, 1, 0)) %>%
  ungroup()

# specify order for plotting
dat.binchoice$Metric = factor(dat.binchoice$Metric, levels=chosen_models)
dat.binchoice$TrialType = factor(dat.binchoice$TrialType, levels=c("AI", "AAN", "AAR"))

if (which_models=="llms_all") {
  dat.binchoice$Category = factor(dat.binchoice$Category, levels=c("human", "GPT2", "BERT", "RoBERTa"))
} else {
  dat.binchoice$Category = factor(dat.binchoice$Category, levels=c("human", "LLMs", "baselines"))
}
```

## Show results

```{r}
# leave only active sentences here
dat.binchoice.active = dat.binchoice %>%
  filter(Voice=="active")
```

### Stats

```{r}
# get p values
dat.binchoice.summary = dat.binchoice.active %>%
  filter(TrialType %in% c("AI", "AAN")) %>%
  group_by(Category, Metric, Model, TrialType) %>%
  summarize(NumCorrect=sum(Accuracy), NumTotal=length(Accuracy)) %>%
  mutate(AccuracyScore = NumCorrect/NumTotal) %>%
  ungroup() %>%
  mutate(pVal = calculate_binom_pval(NumCorrect, NumTotal))

# adjust for multiple comparisons within each category
dat.binchoice.summary = dat.binchoice.summary %>%
  group_by(Category) %>%
  mutate(pValAdjusted = p.adjust(pVal, method="fdr", n=length(pVal)),
         ntoadjust = length(pVal)) %>%
  mutate(pLabel = ifelse(pValAdjusted<0.001, "***", 
                         ifelse(pValAdjusted<0.01, "**",
                                ifelse(pValAdjusted<0.05, "*", "")))) %>%
  ungroup()
```

```{r}
# get human responses separately
human.results = dat.binchoice.summary %>%
  filter(Metric=='human') %>%
  select(TrialType, NumCorrect, NumTotal) %>%
  rename(NumCorrectHuman=NumCorrect, NumTotalHuman=NumTotal)

dat.binchoice.summary.withchisq = merge(dat.binchoice.summary, human.results)
dat.binchoice.summary.withchisq = dat.binchoice.summary.withchisq %>%
  mutate(ChiSq = calculate_chisq_vectorized_chi(NumCorrect, NumTotal, NumCorrectHuman),
         pVal2humans = calculate_chisq_vectorized_p(NumCorrect, NumTotal, NumCorrectHuman)) %>%
  group_by(Category) %>%
  mutate(pVal2humansAdjusted = p.adjust(pVal2humans, method="fdr", n=length(pVal2humans)),
         ntoadjust = length(pVal2humans)) %>%
  mutate(pLabel2humans = ifelse(pVal2humansAdjusted<0.001, "***", 
                         ifelse(pVal2humansAdjusted<0.01, "**",
                                ifelse(pVal2humansAdjusted<0.05, "*", "")))) 

# print the result
for (i in seq_along(dat.binchoice.summary.withchisq$Model)) {
  print(paste(dat.binchoice.summary.withchisq$Model[i], ": ",
        round(dat.binchoice.summary.withchisq$AccuracyScore[i],2), 
        ", χ2=", round(dat.binchoice.summary.withchisq$ChiSq[i],2), 
        ", p=", round(dat.binchoice.summary.withchisq$pVal2humansAdjusted[i],3),
        ";", sep=""))
}
```

### Plot

```{r, echo=FALSE}

plot.binacc = ggplot(data=subset(dat.binchoice.active, TrialType%in%c("AI","AAN")), 
       mapping=aes(x=Metric, y=Accuracy, fill=Category))+
  facet_wrap(~TrialType, ncol = 2)+
  geom_hline(yintercept=1, color='gray50', linetype='dotted')+
  stat_summary(geom='col', fun.y='mean',
               color='black', width=0.8)+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', size = 0.5, width=0.1)+
  geom_text(mapping=aes(x=Metric, y=0.05, label=pLabel), data=dat.binchoice.summary)+
  coord_cartesian(ylim=c(0,1))+
  geom_hline(yintercept=.5, linetype='dotted')+
  theme_classic()+
  labs(x=NULL)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
plot.binacc

savename <- "1_binaryAccuracy_split.png"
ggsave(paste(savedir,savename,sep="/"), width=15, height=12, units='cm')
```

### Stats: difference between AI & AAN conditions

Hypothesis test of proportions (Two proportions z test):
prop.test(c(nr_successes1, nr_successes2), c(nr_total1, nr_total2), p = NULL, alternative = "two.sided",
          correct = TRUE)
          
The function returns:
- the value of Pearson’s chi-squared test statistic.
- a p-value
- a 95% confidence intervals
- an estimated probability of success (the proportion of smokers in the two groups)

```{r, echo=FALSE}
dat.aivsaan = dat.binchoice.summary %>% 
  dplyr::select(TrialType, Metric, NumCorrect, NumTotal)

df_ai.vs.aan = data.frame()

message(paste("Hypothesis test for proportions | AI vs. AAN accuracy"))
for (i in seq_along(chosen_models)) {
  curr_data = subset(dat.aivsaan, Metric==chosen_models[i])
  cnt_successes = curr_data$NumCorrect
  cnt_total = curr_data$NumTotal
  
  proptest <- prop.test(x = cnt_successes, n = cnt_total)
  
  proptest.stat = proptest$statistic
  proptest.pval = proptest$p.value
  
  print(paste(chosen_models[i], "|", proptest.stat, "|", proptest.pval, "|", plabel(proptest.pval)))
  
  df <- data.frame(chosen_models[i], proptest.stat, proptest.pval, plabel(proptest.pval))
  df_ai.vs.aan <- rbind(df_ai.vs.aan,df)
}

colnames(df_ai.vs.aan) = c("Metric", "statistic", "pVal", "pLabel")
```

### Mean model results
```{r}
dat.binchoice.meanbycategory = dat.binchoice.summary %>%
  group_by(Category, TrialType) %>%
  summarize(meanAccuracy = mean(AccuracyScore))
```

## Error analysis

```{r}
acc_by_sentence.LLM = dat.binchoice.active %>% 
  filter(TrialType %in% c("AI", "AAN")) %>%
  filter(Category=='LLMs') %>%
  group_by(ItemNum, Voice) %>%
  mutate(meanAccuracy=mean(Accuracy)) %>% 
  ungroup()
```

### Quick plot
```{r}
ggplot(acc_by_sentence.LLM, aes(x=meanAccuracy)) + geom_histogram() 
```

### Save hard sentences

```{r}
threshold = 0.33

hard_sents = acc_by_sentence.LLM %>%
  filter(meanAccuracy<threshold) %>%
  select(ItemNum, Voice, meanAccuracy)

hard_sents = inner_join(hard_sents, dat) %>%
  select(ItemNum, Sentence, TrialType, Plausibility, meanAccuracy) %>% 
  distinct() %>%
  spread(Plausibility, Sentence)

savename <- "hard_sentences_LLMs.csv"
write.csv(hard_sents, paste(savedir,savename,sep="/"), row.names=FALSE)
```

### Compare with human scores

```{r}
dat.binchoice.human = dat.binchoice %>% filter(Metric=='human')
dat.binchoice.human = inner_join(dat.binchoice.human, acc_by_sentence.LLM %>% select(ItemNum, TrialType, Voice, meanAccuracy)) %>%
  distinct()

# calculate correlation
cor_df = dat.binchoice.human %>% 
  group_by(TrialType) %>%
  summarize(Correlation = cor(meanAccuracy, FinalScoreDiff),
            pVal = cor.test(meanAccuracy, FinalScoreDiff, method="pearson")$p.value) %>%
  mutate(pLabel = ifelse(pVal<0.001, "***",
                               ifelse(pVal<0.01, "**",
                                      ifelse(pVal<0.05, "*", ""))))
  
# plot
ggplot(dat.binchoice.human, aes(x=meanAccuracy, y=FinalScoreDiff)) + 
  stat_summary(geom='col', fun.y='mean', fill='gray80')+
  stat_summary(geom='errorbar', fun.data='mean_se',
               color = 'black', size = 0.7, width=0.07)+
  geom_hline(yintercept=1, color='gray50', linetype='dotted')+
  geom_jitter(width=0.01, size=0.7, color='gray50')+
  facet_wrap(~TrialType, ncol = 2)+
  geom_text(mapping=aes(x=0, y=1.10, 
                      label=sprintf("r = %.2f%s", round(Correlation,2), pLabel)), 
          data=cor_df, size=cortext_size, hjust = 0)+
  scale_x_continuous(breaks=c(0,.33,.66,1), labels=c(0,1,2,3))+
  labs(x="# correct LLM responses", y="Human score difference")+
  theme_classic()

savename <- "LLM correct vs human score diff.png"
ggsave(paste(savedir,savename,sep="/"), width=18, height=12, units='cm')
```






# DETAILED PLOT

## General settings
```{r}
# color scheme
color_plaus = '#1b9e77'
color_voice = '#d95f02'
color_syn = '#7570b3'
cortext_size = 3
```



## HISTOGRAMS PLAUSIBILITY 

```{r, fig.height=2,fig.width=5}
dat.sub = dat %>%
  filter(Voice=="active") %>%
  filter(TrialType%in%c("AI","AAN"))

dat.sub$Metric = factor(dat.sub$Metric, levels=chosen_models)
dat.sub$TrialType = factor(dat.sub$TrialType, levels=c("AI", "AAN"))

plot.plaus.density.sub = ggplot()+
  geom_density(data=dat.sub, mapping=aes(FinalNormScore, fill = factor(Plausibility, levels=c("Implausible","Plausible"))), alpha = 0.2)+
  facet_grid(TrialType~Metric)+
  ggtitle('PLAUSIBILITY')+
  geom_vline(xintercept=reference_value, linetype='dotted')+
  #coord_flip()+
  theme_classic()+
  xlab("Normed score") +
  ylab("Density")+
  theme(plot.title = element_text(hjust=0.5,face='bold'))+
  theme(legend.title = element_blank())
plot.plaus.density.sub
```

### Mean diff scores by category
```{r mean diff, echo=FALSE}
dat.meanDiff = dat.binchoice.active %>% 
  filter(TrialType %in% c("AI", "AAN")) %>%
  group_by(Category, TrialType, Model) %>% 
  summarize(meanDiff = mean(FinalScoreDiff)) %>%
  ungroup() %>%
  group_by(Category, TrialType) %>%
  summarize(meanDiff = round(mean(meanDiff), 2)) %>%
  ungroup()

dat.meanDiff = dat.meanDiff %>% 
  spread(TrialType, meanDiff)
```


## ACTIVE VS PASSIVE

### Prep data

```{r, echo=FALSE}
chosen_models = llms_main

dat.voice = dat %>%
  filter(Experiment=='EventsAdapt') %>%
  filter(Metric %in% chosen_models) %>%
  select(ItemNum, TrialType, Plausibility, Voice, Metric, FinalNormScore) %>%
  group_by(ItemNum, TrialType, Plausibility, Voice, Metric) %>%
  summarize(meanScore = mean(FinalNormScore, na.rm=TRUE)) %>%
  spread(Voice, meanScore)
dat.voice$Metric = factor(dat.voice$Metric, levels=chosen_models)
```

### Calculate correlations

```{r}
df_correlation.voice = get_correlation_df("voice", "human", dat.voice, chosen_models)
```

### Plot

```{r}
df_correlation.voice$Metric = factor(df_correlation.voice$Metric, levels=chosen_models)
dat.voice$Metric = factor(dat.voice$Metric, levels=chosen_models)

plot.voice = ggplot(data=dat.voice)+
  geom_point(mapping=aes(x=active, y=passive), size=0.2, color=color_voice)+
  facet_wrap(~Metric, ncol=ncols, nrow=nrows)+
  geom_abline(slope=1, intercept=0, size=0.2)+
  geom_text(mapping=aes(x=0, y=1.10, 
                        label=sprintf("r = %.2f%s", round(Correlation,2), pVal2zeroLabel)), 
            data=df_correlation.voice, size=cortext_size, hjust = 0)+
  coord_cartesian(ylim=c(0,1.15), xlim=c(0,1))+
  scale_x_continuous(breaks=c(0,0.5,1))+
  scale_y_continuous(breaks=c(0,0.5,1))+
  theme_classic()+
  facet_grid(~Metric)+
  ggtitle('VOICE')+
  xlab("Active")+
  ylab("Passive")+
  theme(plot.title = element_text(hjust=0.5,face='bold'))
plot.voice
```


## SYNONYMOUS SENTENCES

### Prep data

```{r, echo=FALSE}
dat.syn = dat %>%
  filter(!is.na(SynonymPair)) %>%
  filter(Metric %in% chosen_models) %>%
  filter(Voice=="active") %>%
  dplyr::select(-ItemNum) %>%
  group_by(TrialType, NumSyn, SynonymPair, Plausibility, Voice, Metric) %>%
  summarize(meanScore = mean(FinalNormScore, na.rm=TRUE)) %>%
  spread(NumSyn, meanScore)

# remove the sentences that don't have synonyms
dat.syn = dat.syn %>%
  filter(!is.na(Version1)) %>%
  filter(!is.na(Version2))

dat.syn$Metric = factor(dat.syn$Metric, levels=chosen_models)
```

### Calculate correlations

```{r}
df_correlation.syn = get_correlation_df("synonym", "human", dat.syn, chosen_models)
```

### Plot


```{r}
df_correlation.syn$Metric = factor(df_correlation.syn$Metric, levels=chosen_models)
dat.syn$Metric = factor(dat.syn$Metric, levels=chosen_models)

plot.syn = ggplot(data=dat.syn)+
  geom_point(mapping=aes(x=Version1, y=Version2), size=0.2, color=color_syn)+
  #facet_wrap(~Metric, ncol=ncols, nrow=nrows)+
  geom_abline(slope=1, intercept=0, size=0.2)+
  geom_point(mapping=aes(x=Version1, y=Version2), size=0.2, color=color_syn)+
  geom_text(mapping=aes(x=0, y=1.10, 
                        label=sprintf("r = %.2f%s", round(Correlation,2), pVal2zeroLabel)), 
            data=df_correlation.syn, size=cortext_size, hjust = 0)+
  coord_cartesian(ylim=c(0,1.15), xlim=c(0,1))+
  scale_x_continuous(breaks=c(0,0.5,1))+
  scale_y_continuous(breaks=c(0,0.5,1))+
  theme_classic()+
  facet_grid(~Metric)+
  ggtitle('SYNONYMOUS SENTENCES')+
  theme(plot.title = element_text(hjust=0.5,face='bold'))
plot.syn
```


## COMBINE

```{r}
plot.plaus.density.sub / (plot.voice + plot.syn)

savename <- "scatterplot_density_voice_syn.png"
ggsave(paste(savedir,savename,sep="/"), height=14,width=30, units='cm')
```
```{r}
plot.binacc / plot.plaus.density.sub / (plot.voice + plot.syn)

savename <- "general_plot.png"
ggsave(paste(savedir,savename,sep="/"), height=25,width=25, units='cm')
```

#REGRESSION

## Determine whether to include Plausibility by ItemNum random slope
```{r regression, echo=FALSE}
metrics = llms_main
pvals_likelihoodratiotest = c()

for(i in seq_along(metrics)){
  metric = metrics[i]
  dat.metric = dat %>% filter(Metric==metric)
  m1 = lmer(NormScore ~ Plausibility*TrialType*Voice + agent_freq_norm + patient_freq_norm + verb_freq_norm + sentence_freq_norm + (1+Plausibility|ItemNum), 
           data=dat.metric, REML=FALSE, 
           control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e6)))
  m0 = lmer(NormScore ~ Plausibility*TrialType*Voice + agent_freq_norm + patient_freq_norm + verb_freq_norm + sentence_freq_norm + (1|ItemNum), 
           data=dat.metric, REML=FALSE, 
           control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e6)))
  pvals_likelihoodratiotest = c(pvals_likelihoodratiotest, anova(m1, m0)[2,8])
}
print(pvals_likelihoodratiotest)
print(lapply(pvals_likelihoodratiotest, function(x) {x<0.05/length(pvals_likelihoodratiotest)}))    # apply Bonferroni correction
```

Conclusion: the model with the Plausibility x ItemNum slope is a better fit for all key metrics, so will stick to it.

## Stats main

```{r regression, echo=FALSE}
results = data.frame()

for(i in seq_along(metrics)){
  metric = metrics[i]
  dat.metric = dat %>% filter(Metric==metric)
  m = lmer(NormScore ~ Plausibility*TrialType*Voice + agent_freq_norm + patient_freq_norm + verb_freq_norm + sentence_freq_norm + (1+Plausibility|ItemNum), 
           data=dat.metric, REML=FALSE, 
           control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e6)))
 d = data.frame(coef(summary(m)))
 d = d %>% rename(pVal=Pr...t..)
 d$Metric = metric 
 d$Parameter = rownames(d)
 d = d[,c(6,7,1:5)]
 results = rbind(results, d)
}
```

## Correct for MC within LLMs and baselines

```{r adjust pVals, echo=FALSE}
results = results %>%
  mutate(Category = ifelse(Metric%in%llms, "LLMs", "baselines")) %>%
  mutate(Category = ifelse(Metric=="human", "human", Category))

results = results %>%
  group_by(Category, Parameter) %>%
  mutate(pValAdjusted = p.adjust(pVal, method="fdr", n=length(pVal))) %>%
  mutate(pLabel = ifelse(pValAdjusted<0.001, "***", 
                         ifelse(pValAdjusted<0.01, "**",
                                ifelse(pValAdjusted<0.05, "*", "")))) %>%
  mutate(pValFinal = ifelse(pValAdjusted<0.001, "<0.001", as.character(round(pValAdjusted,3)))) %>%
  ungroup()

savename <- "regression_by_metric.csv"
write.csv(results, paste(savedir,savename,sep="/"), row.names=FALSE)
```

## Format

```{r format results, echo=FALSE}
reported_params = c("PlausibilityImplausible:TrialTypeAAR",
                    "PlausibilityImplausible:TrialTypeAI",
                    "VoiceA>P",
                    "agent_freq_norm",
                    "patient_freq_norm",
                    "verb_freq_norm",
                    "sentence_freq_norm",
                    "TrialTypeAAR:VoiceA>P",
                    "TrialTypeAI:VoiceA>P",
                    "PlausibilityImplausible:TrialTypeAAR:VoiceA>P",
                    "PlausibilityImplausible:TrialTypeAI:VoiceA>P")

# restructure 
results.clean = results %>% 
  filter(Parameter %in% reported_params) %>%
  mutate(Value = ifelse(pValAdjusted<0.05, 
                        paste(round(Estimate, 2), pLabel, sep=" "),
                        "")) %>%
  select(Metric, Parameter, Value) %>% 
  spread(Metric, Value)

# order and names
results.clean = results.clean %>%
  mutate(Category = ifelse(Parameter %in% c("PlausibilityImplausible:TrialTypeAAR",
                                            "PlausibilityImplausible:TrialTypeAI"),
                           "Core effects", "Surface-level effects")) %>%
  select(c(Category, Parameter, metrics)) %>%
  mutate(Parameter = factor(Parameter, levels=reported_params)) %>%
  arrange(Parameter) %>%
  mutate(Parameter = recode(Parameter, 
                    "PlausibilityImplausible:TrialTypeAAR"="Plausibility (AAN>control)",
                    "PlausibilityImplausible:TrialTypeAI"="Plausibility (AI>AAN)",
                    "VoiceA>P"="Voice",
                    "agent_freq_norm"="Agent frequency",
                    "patient_freq_norm"="Patient frequency",
                    "verb_freq_norm"="Verb frequency",
                    "sentence_freq_norm"="Avg. word frequency",
                    "TrialTypeAAR:VoiceA>P"="Voice x Sentence (AAN>control)",
                    "TrialTypeAI:VoiceA>P"="Voice x Sentence (AI>AAN)",
                    "PlausibilityImplausible:TrialTypeAAR:VoiceA>P"="Plausibility x Voice x Sentence (AAN>control)",
                    "PlausibilityImplausible:TrialTypeAI:VoiceA>P"="Plausibility x Voice x Sentence (AI>AAN)"))

savename <- "regression_by_metric_formatted.csv"
write.csv(results.clean, paste(savedir,savename,sep="/"), row.names=FALSE)  
```


# SI - MODEL HUMAN COMPARISON

## Calculate human 2 human response correlation 
(each human with the average of all the rest)
```{r}
human_dat.comp = human_dat %>%
  filter(!is.na(Score)) %>%
  mutate(FinalNormScore = normalization(Score)) %>%
  select(WorkerId, ItemNum, TrialType, Plausibility, Voice, Sentence, FinalNormScore)

workerIDs = unique(human_dat.comp$WorkerId)
df_correlation = data.frame()

for (i in seq_along(workerIDs)) {
  if (i%%50==0) {    # just to keep track (n=511 for EventsAdapt)
    print(i)
  }
  human_dat.this = human_dat.comp %>% filter(WorkerId==workerIDs[i])
  human_dat.rest = human_dat.comp %>% filter(WorkerId!=workerIDs[i]) %>%
    filter(ItemNum %in% unique(human_dat.this$ItemNum)) %>%
    group_by(ItemNum, Plausibility, Voice, Sentence) %>% 
    summarize(MeanNormScore=mean(FinalNormScore))
  
  human_dat.side2side = merge(human_dat.this, human_dat.rest,
                              by=c("ItemNum", "Plausibility", "Voice"))
  corval= cor(human_dat.side2side$FinalNormScore, human_dat.side2side$MeanNormScore)
  # add vector to a dataframe
  df <- data.frame(workerIDs[i], corval)
  df_correlation <- rbind(df_correlation,df)
}

human2human_corr = mean(df_correlation$corval)
print(paste("The human2human overall correlation value is", human2human_corr))
```

## Sample a random human response for each item 
And compute the avg of all other responses 

```{r}
set.seed(21)
human_dat.sampled = human_dat.comp %>%
  group_by(ItemNum, TrialType, Plausibility, Voice, Sentence) %>%
  summarize(RandomScore = sample(FinalNormScore,1),
            MeanScoreRest = (sum(FinalNormScore)-RandomScore)/(length(FinalNormScore)-1)) %>%
  ungroup()
```

## Prep data

```{r}
dat.human = dat %>% filter(Metric=="human") %>%
  select(ItemNum, TrialType, Sentence, Plausibility, FinalNormScore)
dat.model = dat %>% filter(Metric!="human") %>%
  select(ItemNum, TrialType, Sentence, Plausibility, FinalNormScore, Metric)
dat.model2human = merge(dat.human, dat.model, 
                        by=c("ItemNum", "TrialType", "Sentence", "Plausibility")) %>%
  rename(HumanScore=FinalNormScore.x, ModelScore=FinalNormScore.y)
```

```{r}
# add human2human data
human_dat.sampled = human_dat.sampled %>%
  select(-Voice) %>%
  filter(Sentence %in% unique(dat.model2human$Sentence)) %>%  # make sure it's the same sents
  rename(HumanScore=MeanScoreRest, ModelScore=RandomScore)    # renaming for convenience 
human_dat.sampled$Metric = 'singleHuman'
```

```{r}
dat.model2human = rbind(dat.model2human, human_dat.sampled)
```

Formatting:

```{r, echo=FALSE, fig.height=12, fig.width=15}
dat.model2human = dat.model2human %>%
  #add color info
  mutate(Category = ifelse(Metric%in%llms, "LLMs", "baselines")) %>%
  mutate(Category = ifelse(Metric=="singleHuman", "human", Category))

# specify order for plotting
metric_names = chosen_models
metric_names[1] = "singleHuman"
dat.model2human$Metric = factor(dat.model2human$Metric, levels=metric_names)
dat.model2human$Category = factor(dat.model2human$Category, levels=c("human", "LLMs", "baselines"))
dat.model2human$TrialType = factor(dat.model2human$TrialType, levels=c("AI", "AAN", "AAR"))
```

## Calculate correlations for the plot

```{r}
trialTypes = c("AI", "AAN", "AAR")
df_correlation = get_correlation_df_tt("model2human", "singleHuman", trialTypes, dat.model2human, metric_names)
```

## Plot

```{r}

dat.model2human$TrialType = factor(dat.model2human$TrialType, levels=c("AI","AAN","AAR"))
df_correlation$Metric = factor(df_correlation$Metric, levels=metric_names)
df_correlation$TrialType = factor(df_correlation$TrialType, levels=c("AI","AAN","AAR"))
df_correlation$Metric = factor(df_correlation$Metric, levels=metric_names)

# text size 
cortext_size = 3

plot.model2human = ggplot(dat=dat.model2human)+  
  geom_point(mapping=aes(x=ModelScore, y=HumanScore, color=Category), 
             size=0.1, position=position_jitter(width=0.01))+
  geom_abline(slope=1, intercept=0, size=0.2)+
  geom_point(mapping=aes(x=ModelScore, y=HumanScore, color=Category), 
             size=0.1, position=position_jitter(width=0.01))+
  geom_text(mapping=aes(x=0.2, y=1.10, 
                        label=sprintf("r = %.2f%s", round(Correlation,2), pVal2zeroLabel)), 
            data=df_correlation, size=cortext_size, hjust = 0)+
  coord_cartesian(ylim=c(0,1.15), xlim=c(-0.15,1))+
  scale_x_continuous(breaks=c(0,0.5,1))+
  scale_y_continuous(breaks=c(0,0.5,1))+
  theme_classic()+
  facet_grid(TrialType~Metric)
plot.model2human

savename <- "model2human_byTrialType_LLMs.png"
ggsave(paste(savedir,savename,sep="/"), height=12,width=24, units='cm')
```

```{r}

dat.model2human.sub = dat.model2human %>% 
  filter(TrialType %in% c("AI","AAN"))

df_correlation = df_correlation %>%
  filter(TrialType %in% c("AI","AAN"))

dat.model2human.sub$TrialType = factor(dat.model2human.sub$TrialType, levels=c("AI","AAN","AAR"))
df_correlation$Metric = factor(df_correlation$Metric, levels=metric_names)
df_correlation$TrialType = factor(df_correlation$TrialType, levels=c("AI","AAN","AAR"))
df_correlation$Metric = factor(df_correlation$Metric, levels=metric_names)

# text size 
cortext_size = 3

plot.model2human.sub = ggplot(dat=dat.model2human.sub)+  
  geom_point(mapping=aes(x=ModelScore, y=HumanScore, color=Category), 
             size=0.1, position=position_jitter(width=0.01))+
  geom_abline(slope=1, intercept=0, size=0.2)+
  geom_point(mapping=aes(x=ModelScore, y=HumanScore, color=Category), 
             size=0.1, position=position_jitter(width=0.01))+
  geom_text(mapping=aes(x=-0.25, y=1, 
                        label=sprintf("r = %.2f%s", round(Correlation,2), pVal2zeroLabel)), 
            data=df_correlation, size=cortext_size, hjust = 0)+
  coord_cartesian(ylim=c(0,1), xlim=c(-0.2,1))+
  theme_classic()+
  facet_grid(TrialType~Metric)
plot.model2human.sub

savename <- "model2human_byTrialType_LLMs_noAAR.png"
ggsave(paste(savedir,savename,sep="/"), height=10,width=25, units='cm')
```

# SI - PLAUSIBLE VS IMPLAUSIBLE

### Prep data
```{r, echo=FALSE}
chosen_models = all_main

dat.plaus = dat %>% 
  dplyr::select(ItemNum, TrialType, Plausibility, Voice, Metric, FinalNormScore) %>%
  group_by(ItemNum, TrialType, Plausibility, Voice, Metric) %>%
  summarize(meanScore = mean(FinalNormScore, na.rm=TRUE)) %>%
  spread(Plausibility, meanScore)  
```

### Calculate correlations

```{r}
trialTypes = c("AI", "AAN")
dat.plaus.sub = dat.plaus %>%
  filter(TrialType%in%c(trialTypes))
df_correlation.plaus.AI = get_correlation_df("plausibility", "human", 
                                                 dat.plaus.sub %>% filter(TrialType=="AI"),
                                                 chosen_models)
df_correlation.plaus.AI$TrialType = "AI"
df_correlation.plaus.AAN = get_correlation_df("plausibility", "human", 
                                                 dat.plaus.sub %>% filter(TrialType=="AAN"),
                                                 chosen_models)
df_correlation.plaus.AAN$TrialType = "AAN"
df_correlation.plaus = rbind(df_correlation.plaus.AI, df_correlation.plaus.AAN)
df_correlation.plaus$TrialType = factor(df_correlation.plaus$TrialType, levels=c("AI", "AAN"))
```

### Plot
```{r}
#determine plotting order
df_correlation.plaus$Metric = factor(df_correlation.plaus$Metric, levels=chosen_models)
dat.plaus.sub$Metric = factor(dat.plaus.sub$Metric, levels=chosen_models)
dat.plaus.sub$TrialType = factor(dat.plaus.sub$TrialType, levels=c("AI", "AAN"))

plot.plaus = ggplot(data=dat.plaus.sub)+
  geom_point(mapping=aes(x=Plausible, y=Implausible), size=0.2, color=color_plaus)+
  facet_grid(TrialType~Metric)+
  geom_abline(slope=1, intercept=0, size=0.2)+
  geom_point(mapping=aes(x=Plausible, y=Implausible), size=0.2, color=color_plaus)+
  geom_text(mapping=aes(x=0, y=1.10, 
                        label=sprintf("r = %.2f%s", round(Correlation,2), pVal2zeroLabel)), 
            data=df_correlation.plaus, size=cortext_size, hjust = 0)+
  coord_cartesian(ylim=c(0,1.15), xlim=c(0,1))+
  scale_x_continuous(breaks=c(0,0.5,1))+
  scale_y_continuous(breaks=c(0,0.5,1))+
  ggtitle('PLAUSIBILITY')+
  theme_classic()+
  theme(plot.title = element_text(hjust=0.5,face='bold'))
plot.plaus
```


